---
title: "Generate Oligo List for MPRA"
output:
  pdf_document: default
  html_document: default
  word_document: default
date: "2023-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generate Bed file for ROI


```{r , warning = FALSE, message = FALSE}
library(dplyr)


#### This script generates oligo sequences for an array from a bed file. 
#### Input is a bedfile with at least 4 columns (chr, start, end, and name)
####
#### EDIT THESE PARAMETERS
####
# set element size and coverage. Sliding window is elementsize/coverage rounded
# down (56.6 = 56)
# give ending file a name
elementsize <- 270
coverage <- 3
slidewindow <- floor(elementsize/coverage)
filename <- "bedfile_nostrand.bed"

####
#### READ IN BEDFILE
####

# assuming we have a bed file with at least 4 columns, chr, start, end, and name
bedfile <-  as_tibble(read.table("gapsandbrilucpos.bed", 
                                 header = FALSE))
bedfile <- bedfile %>% dplyr::rename( "chr" = "V1", 
                                      "start" = "V2", 
                                      "end" = "V3", 
                                      "name" = "V4")



# extend region start and end by elementsize to ensure full coverage from start
# to end of seq
bedextend <- bedfile %>% mutate(start = start - elementsize) %>%
  mutate(end = end + elementsize)

# initialize vectors for for loop
allstart <- c()
allnames <- c()
allchr <- c()

# for each row of bedfile, generate list of sequence "start" values from bed start to end 
# by sliding window size, add to vector with numbered names/chr number
for (i in 1:nrow(bedextend)){
  segstart <- seq(as.numeric(bedextend[i, 2]), 
                  as.numeric(bedextend[i, 3]), 
                  by = slidewindow)
  names <- c(paste(bedextend[i,4], 
                   "_", 
                   1:length(segstart), 
                   sep =""))
  chr <- rep(bedextend[i, 1], length(segstart))
  allstart <- append(allstart, segstart)
  allnames <- append(allnames, names)
  allchr <- append(allchr, chr)
}

# create tibbles with all the chr name, start, end (start + element size), and name
# duplicate tibble, add columns for + and - strands
# merge tibbles so we have each segment, + and - strands
tib <- tibble(chr = as.character(allchr), 
              start = allstart, 
              end = allstart + elementsize, 
              name = allnames)
tib$score <-  100
#tib2 <- tib
#tib$strand <- "+"
#tib2$strand <- "-"
#tib$name <- paste(tib$name, 
#                  "_+", 
 #                 sep = "")
#tib2$name <- paste(tib2$name, 
  #                 "_-", 
   #                sep = "")
#tibmerge <- bind_rows(tib, tib2)

tib

# write all beds to a bedfile
write.table(tib, 
            file = filename, 
            sep = "\t", 
            col.names = FALSE, 
            row.names = FALSE, 
            quote = FALSE)

```



# send output to bash script below
generates a tab separated list of sequences from the generated bedfile, change the reference genome to the appropriate genome

# bash script
module load cluster/bedtools/2.28.0

bedtools sort -i bedfile.bed > sort.bedfile.bed

bedtools getfasta -tab -name+ -s -fi /cluster/home/rhauser/genomes/chr17.fa -bed sort.bedfile.bed > fastalist.tsv

# import output to r script below




```{r addadapters, warning = FALSE, message = FALSE}
####
#### READ IN TAB SEPARATED FASTA FILE
#### SET NAME FOR OUTPUT FASTA FILE
####
dat <- as_tibble(read.table(file = "fastalist.tsv", 
                            sep ="\t", 
                            header = FALSE))
filename2 <- "finalfastalist.fa"

# rename columns to make sense
dat <- dat %>% dplyr::rename( "name" = "V1", 
                              "sequence" = "V2")


# add ">" to sequence name to prep for fasta format
dat <-   mutate(dat, name = paste0(">", name))

# add adapters to sequences (from Gordon et al 2020,
# https://doi.org/10.1038/s41596-020-0333-5)
dat <-  mutate(dat, sequence = paste("AGGACCGGATCAACT", 
                                     sequence, 
                                     "CATTGCGTGAACCGA", 
                                     sep = ""))

dat

# write table of sequences to a fasta file
write.table(dat, 
            file = filename2, 
            sep = "\n", 
            col.names = FALSE, 
            row.names = FALSE, 
            quote = FALSE)

```


## Negative Controls

```{r scram, warning = FALSE, message = FALSE}
# This scripts generates random DNA sequences
library(Biostrings)

# set seed so random seqs reproducible, initialize scram
set.seed(123)
scram <- NULL

####
#### EDIT THESE PARAMETERS
####
# set number of scrambled seq to generate and length of scrambled seq
# set file name for end file
numseq <- 100
seqlen <- 270
filename3 <- "scrambled.fa"

# generates vector of desired number of random DNA seqs (A,C,G,T)
for (i in 1:numseq)
  scram = c(scram, paste(sample(DNA_ALPHABET[1:4], 
                                seqlen, 
                                replace = TRUE), 
                         collapse =""))

# create vector for scrambled DNA seq names
scramnames <- c(paste0("scrambled", 1:numseq))
scrambled <- tibble(scramnames, scram)

# prep for fasta format (add > to names) and add adapters to sequences(from Gordon et al 2020)
scrambled <-   mutate(scrambled, 
                      scramnames = paste0(">", scramnames))
scrambled <-  mutate(scrambled, 
                     scram = paste("AGGACCGGATCAACT", 
                                   scram, "CATTGCGTGAACCGA", 
                                   sep = ""))

scrambled
write.table(scrambled, 
            file = filename3, 
            sep = "\n", 
            col.names = FALSE, 
            row.names = FALSE, 
            quote = FALSE)

```

# to combine ROI and Scrambled DNA FASTA files, use this shell script

#cat finalfastalist.fa scrambled.fa > merged.fa

```{r , warning = FALSE, message = FALSE}
library(dplyr)
library(tidyverse)

#import datasets for areas of interest
justbeds <- as_tibble(read_tsv("Links_GeneList_for_Becky_230209.bed", 
                               col_names = c("chr", "start", "end", "gene", "group")))

library(readxl)
X2023_01_17_E280A_AAO_Revised_SuppTables_2_4_Cochran <- 
  read_excel("2023.01.17_E280A-AAO_Revised-SuppTables_2-4-Cochran.xlsx", 
             sheet = "Supp. Table 3")

e280ajustbeds <- X2023_01_17_E280A_AAO_Revised_SuppTables_2_4_Cochran %>% 
  dplyr::select(seqnames, start, end, gene, group)

e280a_nodups <- e280ajustbeds %>% distinct(seqnames, start, end, .keep_all = TRUE)

```

```{r}

#filtering for just genes we want
app <- filter(justbeds, gene =="APP")
snca <- filter(justbeds, gene=="SNCA")
sorl1 <- filter(justbeds, gene == "SORL1")
sorl1ad <- filter(justbeds, gene == "SORL1", group == "AD")
sorl1com <- filter(justbeds, gene =="SORL1", group =="common")

e280a <- e280a_nodups %>%
  unite("name", gene:group, sep = "_", remove = TRUE) %>% dplyr::rename("chr" = seqnames)

e280a$name <- paste("e280a_", e280a$name, sep ="")

appmerge <- app %>% unite("name", gene:group, sep = "_", remove = TRUE) %>%
  distinct(chr, start, end, .keep_all = TRUE) 

snca <- snca %>% unite("name", gene:group, sep = "_", remove = TRUE) %>%
  distinct(chr, start, end, .keep_all = TRUE)

sorl1 <- sorl1 %>% unite("name", gene:group, sep = "_", remove = TRUE) %>%
  distinct(chr, start, end, .keep_all = TRUE)

sorl1ad <- sorl1ad %>% unite("name", gene:group, sep = "_", remove = TRUE) %>%
  distinct(chr, start, end, .keep_all = TRUE)

sorl1com <- sorl1com %>% unite("name", gene:group, sep = "_", remove = TRUE) %>%
  distinct(chr, start, end, .keep_all = TRUE)

all <- bind_rows(e280a, appmerge, snca, sorl1ad, sorl1com)

all

```

```{r}

#### This script generates oligo sequences for an array from a bed file. 
#### Input is a bedfile with at least 4 columns (chr, start, end, and name)
####
#### EDIT THESE PARAMETERS
####
# set element size and coverage. Sliding window is elementsize/coverage rounded
# down (56.6 = 56)
# give ending file a name
elementsize <- 270
coverage <- 3
slidewindow <- floor(elementsize/coverage)
#filename <- "bedfile.bed"

####
#### READ IN BEDFILE
####

# assuming we have a bed file with at least 4 columns, chr, start, end, and name
#bedfile <-  as_tibble(read.table("gapsandbrilucpos.bed", 
#                                 header = FALSE))
#bedfile <- bedfile %>% dplyr::rename( "chr" = "V1", 
#                                      "start" = "V2", 
#                                      "end" = "V3", 
#                                      "name" = "V4")

bedfile <- all

# extend region start and end by elementsize to ensure full coverage from start
# to end of seq
bedextend <- bedfile %>% mutate(start = start - elementsize) %>%
  mutate(end = end + elementsize)

# initialize vectors for for loop
allstart <- c()
allnames <- c()
allchr <- c()

# for each row of bedfile, generate list of sequence "start" values from bed start to end 
# by sliding window size, add to vector with numbered names/chr number
for (i in 1:nrow(bedextend)){
  segstart <- seq(as.numeric(bedextend[i, 2]), 
                  as.numeric(bedextend[i, 3]), 
                  by = slidewindow)
  names <- c(paste(bedextend[i,4], 
                   "_", 
                   1:length(segstart), 
                   sep =""))
  chr <- rep(bedextend[i, 1], length(segstart))
  allstart <- append(allstart, segstart)
  allnames <- append(allnames, names)
  allchr <- append(allchr, chr)
}

# create tibbles with all the chr name, start, end (start + element size), and name
# duplicate tibble, add columns for + and - strands
# merge tibbles so we have each segment, + and - strands
tib <- tibble(chr = as.character(allchr), 
              start = allstart, 
              end = allstart + elementsize, 
              name = allnames)
tib$score <- 100
tib2 <- tib
tib$strand <- "+"
tib2$strand <- "-"
tib$name <- paste(tib$name, 
                  "_+", 
                  sep = "")
tib2$name <- paste(tib2$name, 
                   "_-", 
                   sep = "")
tibmerge <- bind_rows(tib, tib2)

tibmerge

# write all beds to a bedfile
write.table(tibmerge, 
            file = "addons.bed", 
            sep = "\t", 
            col.names = FALSE, 
            row.names = FALSE, 
            quote = FALSE)

```



# send output to bash script below
generates a tab separated list of sequences from the generated bedfile, change the reference genome to the appropriate genome

# bash script
module load cluster/bedtools/2.28.0

bedtools sort -i addons.bed > sort.addons.bed

bedtools getfasta -tab -s -name+ -fi /cluster/home/rhauser/genomes/hg38/hg38.fa -bed sort.addons.bed > addonfastalist.tsv

# import output to r script below




```{r addadapters, warning = FALSE, message = FALSE}
####
#### READ IN TAB SEPARATED FASTA FILE
#### SET NAME FOR OUTPUT FASTA FILE
####
dat <- as_tibble(read.table(file = "addonfastalist.tsv", 
                            sep ="\t", 
                            header = FALSE))
filename2 <- "addonfinalfastalist.fa"

# rename columns to make sense
dat <- dat %>% dplyr::rename( "name" = "V1", 
                              "sequence" = "V2")


# add ">" to sequence name to prep for fasta format
dat <-   mutate(dat, name = paste0(">", name))

# add adapters to sequences (from Gordon et al 2020,
# https://doi.org/10.1038/s41596-020-0333-5)
dat <-  mutate(dat, sequence = paste("AGGACCGGATCAACT", 
                                     sequence, 
                                     "CATTGCGTGAACCGA", 
                                     sep = ""))

dat

# write table of sequences to a fasta file
write.table(dat, 
            file = filename2, 
            sep = "\n", 
            col.names = FALSE, 
            row.names = FALSE, 
            quote = FALSE)

```


# to combine new ROI and original FASTA list, use this shell script

#cat addonfinalfastalist.fa merged.fa > newmerged.fa
