---
title: "Untitled"
author: "Becky"
date: "2026-03-19"
output: html_document
---
#libraries
```{r}
library(tidyverse)
library(ggplot2)
library(rtracklayer)
library(plotgardener)
library(GenomicRanges)
library(org.Hs.eg.db)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(Biostrings)
library(BSgenome.Hsapiens.UCSC.hg38)
library(DESeq2)
library(edgeR)
library(ggrepel)
library(Seurat)
library(ComplexHeatmap)
library(tidyHeatmap)
library(reshape2)
library(colorspace)
library(circlize)
```

## MPRA output
### import data
```{r}
#import Neurons
bacmpra <- read_tsv("/cluster/home/rhauser/files/rmhex01/MPRAflowCount_3reps/output_mpranalyze/Neurons/testEmpirical_alpha_RMHex01_Neurons_results_labeled_analysis2.tsv")

#extract start location
bacmpra <- bacmpra %>%
  mutate(start = as.integer(str_extract(label, "(?<=bin)\\d+")))
bacmpra$end <- bacmpra$start + 99
bacmpra$chr <- "chr17"
bacmprahits <- bacmpra %>% filter(pval.mad<=0.05)
bacmpra$score <- -log10(bacmpra$pval.mad)
bacmprasignal <- bacmpra %>% dplyr::select(chr, start, end, score)
bacmprasignal <- bacmprasignal %>% arrange(by=start)

gr_bacmpra <- GRanges(seqnames = bacmprahits$chr,
                      ranges = IRanges(start = bacmprahits$start,
                                       end = bacmprahits$end))
gr_bacmpra_reduce <- reduce(gr_bacmpra)


#import HEKs
hekbacmpra <- read_tsv("/cluster/home/rhauser/files/rmhex01/MPRAflowCount_3reps/output_mpranalyze/HEK293FT/testEmpirical_alpha_RMHex01_HEK293FT_results_labeled_analysis2.tsv")
#extract start location
hekbacmpra <- hekbacmpra %>%
  mutate(start = as.integer(str_extract(label, "(?<=bin)\\d+")))
hekbacmpra$end <- hekbacmpra$start + 99
hekbacmpra$chr <- "chr17"

hekbacmprahits <- hekbacmpra %>% filter(pval.mad<=0.05)

gr_hekbacmpra <- GRanges(seqnames = hekbacmprahits$chr,
                      ranges = IRanges(start = hekbacmprahits$start,
                                       end = hekbacmprahits$end))
gr_hekbacmpra_reduce <- reduce(gr_hekbacmpra)

#join bacs and neurons into 1 tibble
bacmpra$group <- "neurons"
hekbacmpra$group <- "HEKs"

hekbacmpratemp <- hekbacmpra %>% dplyr::rename_with(~paste0("hek_", .), -label)
bacboth <- bacmpra %>% left_join(hekbacmpratemp, by="label")
bacboth <- bacboth %>% drop_na()
bacboth <- bacboth %>%
  mutate(group = case_when(
    hek_pval.mad <= 0.05 & pval.mad <= 0.05 ~ "sigboth",
    hek_pval.mad <= 0.05 ~ "sighek",
    pval.mad <= 0.05 ~ "signeuron",
    TRUE ~ "ns"
  ))
```

#### correlation plot of HEK and Neuron Hits
```{r}
baccorrplot <- ggplot(bacboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1.5, alpha=0.4) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c( "#66298E", "grey55", "#C873EF", "#C873EF")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_minimal() +
  theme(legend.position = "none")
  #xlim(0, 15) +   # Set x-axis limits from 0 to 15
  #ylim(0, 15)
ggsave(filename="bacmpra_celltypecorr_pval_mad.pdf", plot=baccorrplot, width=4.44, height=4.44, units="in")

baccorrplotmad <- ggplot(bacboth, aes(x=mad.score, y=hek_mad.score, color = group)) +
  geom_point(size = 1.5, alpha=0.4) +
  geom_vline(xintercept = 1.65, linetype = "dotted", color = "black") +  
  geom_hline(yintercept = 1.65, linetype = "dotted", color = "black") +
 scale_color_manual(values = c( "#66298E", "grey55", "#C873EF", "#C873EF")) +
  labs(x = "Neurons mad.score", y = "HEK293FT mad.score", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_minimal() +
  theme(legend.position = "none")
  #xlim(0, 15) +   # Set x-axis limits from 0 to 15
  #ylim(0, 15)
ggsave(filename="bacmpra_celltypecorr_madscore.pdf", plot=baccorrplotmad, width=4.44, height=4.44, units="in")

baccorrplotcutoff <- ggplot(bacboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1.5, alpha=0.4) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c( "#66298E", "grey55", "#C873EF", "#C873EF")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_minimal() +
  theme(legend.position = "none") +
  xlim(0, 15) +   
  ylim(0, 15)
ggsave(filename="bacmpra_celltypecorr_pval_mad_cutoff.pdf", plot=baccorrplotcutoff, width=4.44, height=4.44, units="in")

#testbac <- bacboth %>% dplyr::select(mad.score, hek_mad.score, group)
#testbac %>% filter(group=="sighek") %>% arrange(by=-hek_mad.score)


baccorrplot <- ggplot(bacboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c( "#66298E", "grey55", "#C873EF", "#C873EF")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(expand = c(0, 0), limits=c(0,15)) +
  scale_y_continuous(expand = c(0, 0), limits=c(0,15))
ggsave(filename="bacmpra_celltypecorr_pval_mad2.pdf", plot=baccorrplot, width=4.44, height=4.44, units="in")

baccorrplot <- ggplot(bacboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c( "#66298E", "grey55", "#C873EF", "#C873EF")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(expand = c(0, 0), limits=c(0,40)) +
  scale_y_continuous(expand = c(0, 0), limits=c(0,40))
ggsave(filename="bacmpra_celltypecorr_pval_mad2_uncropped.pdf", plot=baccorrplot, width=4.44, height=4.44, units="in")

baccorrplot <- ggplot(bacboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c( "#66298E", "grey55", "#C873EF", "#C873EF")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(expand = c(0, 0), limits=c(0,10)) +
  scale_y_continuous(expand = c(0, 0), limits=c(0,40))
ggsave(filename="bacmpra_celltypecorr_pval_mad2_uncropped_notsamescale.pdf", plot=baccorrplot, width=4.44, height=4.44, units="in")

bacboth <- bacboth %>%
  mutate(pval.mad = if_else(pval.mad < 2.2e-16, 2.2e-16, pval.mad))
bacboth <- bacboth %>%
  mutate(hek_pval.mad = if_else(hek_pval.mad < 2.2e-16, 2.2e-16, hek_pval.mad))
baccorrplot <- ggplot(bacboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c( "#66298E", "grey55", "#C873EF", "#C873EF")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(expand = c(0, 0), limits=c(0,16)) +
  scale_y_continuous(expand = c(0, 0), limits=c(0,16))
ggsave(filename="bacmpra_celltypecorr_pval_mad2_uncropped_artlimitpval.pdf", plot=baccorrplot, width=4.44, height=4.44, units="in")
```

#### barplots of total hits per cell type
```{r}
#make count tibble
bactotal <- tibble(
  id = c("neuronhits", "neuronreduce", "hekhits", "hekreduce"),
  celltype = c("neurons", "neurons", "heks", "heks"),
  group = c("all", "reduced", "all", "reduced"),
  count = c(length(gr_bacmpra), length(gr_bacmpra_reduce), length(gr_hekbacmpra), length(gr_hekbacmpra_reduce))
)

# Set group order
bactotal$celltype <- factor(bactotal$celltype, levels = c("neurons", "heks"))

bactotalbar <- ggplot(bactotal, aes(x=celltype, y=count, fill=group)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", color="#C873EF", width=0.7) +
  scale_fill_manual(values=c("#C873EF", "white")) +
  theme_minimal()
ggsave(filename="bactotalhits.pdf", plot=bactotalbar, width=4.44, height=4.44, units="in")

```

## overlap of ENCODE CREs
```{r}
encodecres_gr <- GRanges(encodecres)
priority_levels <- c("PLS", "pELS", "dELS", "CA", "CA-CTCF","CA-H3K4me3", "CA-TF", "TF", "None")

gr_hekbacmpra_reduce$creType <- "None"
overlapheks <- findOverlaps(gr_hekbacmpra_reduce, encodecres_gr)
annot_df <- data.frame(peak_index = queryHits(overlapheks), annotation = encodecres_gr$creType[subjectHits(overlapheks)])
annot_df$priority <- match(annot_df$annotation, priority_levels)
final_annot <- annot_df %>% group_by(peak_index) %>% slice_min(order_by=priority, n=1) %>% ungroup()
gr_hekbacmpra_reduce$creType[final_annot$peak_index] <- final_annot$annotation

gr_bacmpra_reduce$creType <- "None"
overlapneur <- findOverlaps(gr_bacmpra_reduce, encodecres_gr)
annot_df_neur <- data.frame(peak_index = queryHits(overlapneur), annotation = encodecres_gr$creType[subjectHits(overlapneur)])
annot_df_neur$priority <- match(annot_df_neur$annotation, priority_levels)
final_annot_neur <- annot_df_neur %>% group_by(peak_index) %>% slice_min(order_by=priority, n=1) %>% ungroup()
gr_bacmpra_reduce$creType[final_annot_neur$peak_index] <- final_annot_neur$annotation


gr_hekhitstib <- as_tibble(gr_hekbacmpra_reduce)
gr_hekhitstib$group <- "HEK293FT"
gr_neuronhitstib <- as_tibble(gr_bacmpra_reduce)
gr_neuronhitstib$group <- "Neurons"

counttib_neurons <- gr_neuronhitstib %>%
  dplyr::count(creType, name = "count")
counttib_neurons <- counttib_neurons %>%
  mutate(fraction = count / sum(count))
counttib_neurons$group <- "Neurons"

counttib_heks <- gr_hekhitstib %>% dplyr::count(creType, name="count")
counttib_heks <- counttib_heks %>% mutate(fraction = count / sum(count))
counttib_heks$group <- "HEK293FT"

results_oligocombined <- rbind(counttib_heks, counttib_neurons)

# Set group order
results_oligocombined$group <- factor(results_oligocombined$group, levels = c("Neurons", "HEK293FT"))

hekneuronoligocrestack_fraction <- ggplot(results_oligocombined, aes(x=group, y=fraction, fill = fct_reorder(creType, fraction))) + 
  geom_bar(position="stack", stat="identity", show.legend = TRUE) +
  theme_minimal() + 
  ylab("Percent CRE Overlap") + 
  ggtitle("Overlap Count by CRE Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_classic() +
  scale_fill_manual(values = color_list) +
  labs(fill = "Encode CREs") +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
ggsave(filename="creoverlap_reducedhits_bacmpra_heksandneurons_stackedbar_fraction4_reducedhitsashlynway.pdf", plot=hekneuronoligocrestack_fraction, width=4.44, height=4.44, units="in")


```

### import data - Neurons (one tailed test)
```{r}
#load neuron hits with scrambled
neurons <- read_tsv("/cluster/home/rhauser/files/rmhex02/MPRAflowCount_reps123/output_mpranalyze/Neurons/testEmpirical_alpha_Neurons_results_labeled_prefilt_allcontrols_oneside.tsv")

neurons$neglog10pval <- -log10(neurons$pval.mad)
neurons_noscram <- neurons %>% filter(!str_detect(label, "scram"))
neurons_noscram$label2 <- neurons_noscram$label
#split up label column into usable chr, start, end info
neurons_noscram <- neurons_noscram %>% separate(label2, into = c("prefix", "chr", "pos"), sep = "::") 
neurons_noscram <- neurons_noscram %>% separate(chr, into = c("chr", "start"), sep = ":")
neurons_noscram<- neurons_noscram %>% separate(start, into = c("start", "end"), sep = "-")
neurons_noscram <- neurons_noscram %>% mutate(end = str_remove(end, "\\+$"))
#neurons_noscram <- neurons_noscram %>% dplyr::select(chr, start, end, mad.score, pval.mad, neglog10pval, label)
neurons_noscram <- neurons_noscram %>% arrange(by=start)
neurons_noscram$start <- as.integer(neurons_noscram$start)
neurons_noscram$end <- as.integer(neurons_noscram$end)

#neurons_noscram <- neurons_noscram %>% dplyr::select(chr, start, end, neglog10pval)
neurons_hits <- neurons_noscram %>% dplyr::select(chr, start, end, pval.mad)
neurons_hits <- neurons_hits %>% filter(pval.mad <=0.05)
neurons_hits <- neurons_hits %>% arrange(by=start)
neurons_hits01 <- neurons_noscram %>% dplyr::select(chr,start,end, pval.mad) %>%
  filter(pval.mad <= 0.01) %>%
  arrange(by=start)


gr_neuronhits <- GRanges(seqnames = neurons_hits$chr,
                      ranges = IRanges(start = as.numeric(neurons_hits$start),
                                       end = as.numeric(neurons_hits$end)))
gr_neuronhits_reduce <- GenomicRanges::reduce(gr_neuronhits)




#mean signal
avg_neurons <- read_tsv("/cluster/home/rhauser/files/rmhex02/onesidedneurons3repswithcontrolsneglog10pval.sorted.mean.bed", col_names=c("chr", "start", "end", "neglog10pval"))
avg_neurons$start <- avg_neurons$start + 1
avg_neurons$score <- avg_neurons$neglog10pval

avg_neurons <- avg_neurons %>% arrange(by=start)
```

## import data -HEKs (one tailed test)
```{r}
#HEKs
#get rmhex02 hek hits, 3 reps, one sided test with negative controls
heks <- read_tsv("/cluster/home/rhauser/files/rmhex02/MPRAflowCount_reps123/output_mpranalyze/HEK293FT/testEmpirical_alpha_HEK293FT_results_labeled_prefilt_allcontrols.tsv")

#remove scrambled sequences, if any were positive hits
heks_noscram <- heks %>% filter(!str_detect(label, "scram"))
#split up label column into usable chr, start, end info
heks_noscram$label2 <- heks_noscram$label
heks_noscram <- heks_noscram %>% separate(label2, into = c("prefix", "chr", "pos"), sep = "::") 
heks_noscram <- heks_noscram %>% separate(chr, into = c("chr", "start"), sep = ":")
heks_noscram <- heks_noscram %>% separate(start, into = c("start", "end"), sep = "-")
heks_noscram <- heks_noscram %>% mutate(end = str_remove(end, "\\+$"))

hek_hits <- heks_noscram %>% filter(pval.mad<=0.05)
gr_hekhits <- GRanges(seqnames = hek_hits$chr,
                      ranges = IRanges(start = as.numeric(hek_hits$start),
                                       end = as.numeric(hek_hits$end)))
gr_hekhits_reduce <- GenomicRanges::reduce(gr_hekhits)
```

#### combine hek and neuron data
```{r}
hekoligompratemp <- heks_noscram %>% dplyr::rename_with(~paste0("hek_", .), -label)
oligoboth <- neurons_noscram %>% left_join(hekoligompratemp, by="label")
oligoboth <- oligoboth %>%
  mutate(group = case_when(
    hek_pval.mad <= 0.05 & pval.mad <= 0.05 ~ "sigboth",
    hek_pval.mad <= 0.05 ~ "sighek",
    pval.mad <= 0.05 ~ "signeuron",
    TRUE ~ "ns"
  ))
#oligoboth <- oligoboth %>% drop_na()


```

# import other data
```{r}
bricrisprpos_noprom <- tibble(
  chr="chr17",
  start=c(45219248, 45241086, 45309310, 45428887, 45431544, 45848403, 45873588, 45940785, 45942182, 45971524),
  end=c(45219836, 45242238, 45309777, 45429757, 45432556, 45849786, 45874056, 45941252, 45942650, 45971991),
  name=c(-674458, -652338, -584456, -464677, -461949, -44905, -20178, 47019, 48416, 77758)
)
bricrisprpos_noprom$size <- bricrisprpos_noprom$end - bricrisprpos_noprom$start
bricrisprpos_noprom$start <- as.numeric(bricrisprpos_noprom$start)
bricrisprpos_noprom$end <- as.numeric(bricrisprpos_noprom$end)

variants1 <- read_tsv("/cluster/home/rhauser/files/rmhex02/ADSP_chr17_variation_mpra_overlap_9-5-24.list.tsv", skip=3466)
variants1bed <- variants1 %>% dplyr::select('#CHROM', POS)
variants1bed$end <- variants1$POS + 1
variants1 <- variants1 %>% arrange(by=POS)

mapt_snmultiomics_brain <- read_tsv("/cluster/home/rhauser/files/rogers2024supplementaldata/mapt_multiomics_brain.bed", skip=1, col_names=c("chr","start","end"))

maptluc <- read_tsv("/cluster/home/rhauser/files/rogers2024supplementaldata/luciferaseregions.bed", skip=1, col_names=c("chr","start","end"))

maptcrispr <- read_tsv("/cluster/home/rhauser/files/rogers2024supplementaldata/CRISPRguides.bed", skip=1, col_names=c("chr", "start", "end"))

maptcapturec_gluta <- read_tsv("/cluster/home/rhauser/files/rogers2024supplementaldata/mapt_gluta_capturec.bed", skip=1, col_names=c("chr","start","end"))

mapt_snmultiomics_culture <- read_tsv("/cluster/home/rhauser/files/rogers2024supplementaldata/mapt_multiomics_neurons.bed", skip=1, col_names=c("chr", "start", "end"))

mapt_hicgluta <- read_tsv("/cluster/home/rhauser/files/rogers2024supplementaldata/mapt_hic_hiccups_gluta.bed", skip=1, col_names=c("chr", "start", "end"))

maptcapturec_gaba <- read_tsv("/cluster/home/rhauser/files/rogers2024supplementaldata/mapt_gaba_capturec.bed", skip=1, col_names=c("chr", "start", "end"))

beckycrispr <- read_tsv("/cluster/home/rhauser/files/rmhex02/CRISPRi/rmhex02crisprguidesfinal.bed", skip=1, col_names=c("chr", "start", "end", "region"))
```

### overlap of ENCODE CREs
```{r}
encodecres_gr <- GRanges(encodecres)
priority_levels <- c("PLS", "pELS", "dELS", "CA", "CA-CTCF","CA-H3K4me3", "CA-TF", "TF", "None")

gr_hekhits_reduce$creType <- "None"
overlapheks <- findOverlaps(gr_hekhits_reduce, encodecres_gr)
annot_df <- data.frame(peak_index = queryHits(overlapheks), annotation = encodecres_gr$creType[subjectHits(overlapheks)])
annot_df$priority <- match(annot_df$annotation, priority_levels)
final_annot <- annot_df %>% group_by(peak_index) %>% slice_min(order_by=priority, n=1) %>% ungroup()
gr_hekhits_reduce$creType[final_annot$peak_index] <- final_annot$annotation

gr_neuronhits_reduce$creType <- "None"
overlapneur <- findOverlaps(gr_neuronhits_reduce, encodecres_gr)
annot_df_neur <- data.frame(peak_index = queryHits(overlapneur), annotation = encodecres_gr$creType[subjectHits(overlapneur)])
annot_df_neur$priority <- match(annot_df_neur$annotation, priority_levels)
final_annot_neur <- annot_df_neur %>% group_by(peak_index) %>% slice_min(order_by=priority, n=1) %>% ungroup()
gr_neuronhits_reduce$creType[final_annot_neur$peak_index] <- final_annot_neur$annotation


gr_hekhitstib <- as_tibble(gr_hekhits_reduce)
gr_hekhitstib$group <- "HEK293FT"
gr_neuronhitstib <- as_tibble(gr_neuronhits_reduce)
gr_neuronhitstib$group <- "Neurons"

counttib_neurons <- gr_neuronhitstib %>%
  dplyr::count(creType, name = "count")
counttib_neurons <- counttib_neurons %>%
  mutate(fraction = count / sum(count))
counttib_neurons$group <- "Neurons"

counttib_heks <- gr_hekhitstib %>% dplyr::count(creType, name="count")
counttib_heks <- counttib_heks %>% mutate(fraction = count / sum(count))
counttib_heks$group <- "HEK293FT"

results_oligocombined <- rbind(counttib_heks, counttib_neurons)

# Set group order
results_oligocombined$group <- factor(results_oligocombined$group, levels = c("Neurons", "HEK293FT"))

hekneuronoligocrestack_fraction <- ggplot(results_oligocombined, aes(x=group, y=fraction, fill = fct_reorder(creType, fraction))) + 
  geom_bar(position="stack", stat="identity", show.legend = TRUE) +
  theme_minimal() + 
  ylab("Percent CRE Overlap") + 
  ggtitle("Overlap Count by CRE Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_classic() +
  scale_fill_manual(values = color_list) +
  labs(fill = "Encode CREs") +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
ggsave(filename="creoverlap_reducedhits_oligompra_heksandneurons_stackedbar_fraction4_reducedhitsashlynway.pdf", plot=hekneuronoligocrestack_fraction, width=4.44, height=4.44, units="in")

```


#### correlation plot of Neurons and HEKs hits
```{r}
oligocellcorr <- ggplot(oligoboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1.5, alpha=0.4) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c("#27706E","gray55", "#4FE0DC", "#4FE0DC")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_minimal() +
  theme(legend.position = "none")
ggsave(filename="oligompra_celltypecorr_pval_mad.pdf", plot=oligocellcorr, width=4.44, height=4.44, units="in")

oligocellcorrmad <- ggplot(oligoboth, aes(x=mad.score, y=hek_mad.score, color = group)) +
  geom_point(size = 1.5, alpha=0.4) +
  #geom_vline(xintercept = 1.6, linetype = "dotted", color = "black") +  
  #geom_hline(yintercept = 1.96, linetype = "dotted", color = "black") +
 scale_color_manual(values = c("#027C89","gray55", "#4FE0DC", "#4FE0DC")) +
  labs(x = "Neurons mad.score", y = "HEK293FT mad.score", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_minimal() +
  theme(legend.position = "none")
ggsave(filename="oligompra_celltypecorr_madscore_nodottedline.pdf", plot=oligocellcorrmad, width=4.44, height=4.44, units="in")

oligobothtest <- oligoboth %>% dplyr::select(label, group, mad.score, hek_mad.score)
oligobothtest <- oligobothtest %>% filter(group=="signeuron")
oligobothtest %>% arrange(by=-mad.score)
summary(oligobothtest)

oligocellcorrcutoff <- ggplot(oligoboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1.5, alpha=0.4) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c("#27706E","gray55", "#4FE0DC", "#4FE0DC")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_minimal() +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0, 10)) +
  scale_y_continuous(limits = c(0, 10))
ggsave(filename="oligompra_celltypecorr_pval_cutoff.pdf", plot=oligocellcorrcutoff, width=4.44, height=4.44, units="in")

oligoboth <- oligoboth %>%
  mutate(pval.mad = if_else(pval.mad < 2.2e-16, 2.2e-16, pval.mad))
oligoboth <- oligoboth %>%
  mutate(hek_pval.mad = if_else(hek_pval.mad < 2.2e-16, 2.2e-16, hek_pval.mad))
oligocellcorrcutoff <- ggplot(oligoboth, aes(x=-log10(pval.mad), y=-log10(hek_pval.mad), color = group)) +
  geom_point(size = 1) +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "black") +
 scale_color_manual(values = c("#27706E","gray55", "#4FE0DC", "#4FE0DC")) +
  labs(x = "Neurons -log10(pval.mad)", y = "HEK293FT -log10(pval.mad)", title = "MPRA Correlation\nNeurons vs. HEK293FT") + 
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(expand = c(0, 0), limits=c(0,16)) +
  scale_y_continuous(expand = c(0, 0), limits=c(0,16))
ggsave(filename="oligompra_celltypecorr_pval_mad2_uncropped_artlimitpval.pdf", plot=oligocellcorrcutoff, width=4.44, height=4.44, units="in")

```
#### barplots of total hits per cell type
```{r}
#make count tibble
oligototal <- tibble(
  id = c("neuronhits", "neuronreduce", "hekhits", "hekreduce"),
  celltype = c("neurons", "neurons", "heks", "heks"),
  group = c("all", "reduced", "all", "reduced"),
  count = c(length(gr_neuronhits), length(gr_neuronhits_reduce), length(gr_hekhits), length(gr_hekhits_reduce))
)

#set order
oligototal$celltype <- factor(oligototal$celltype, levels = c("neurons", "heks"))

oligototalbar <- ggplot(oligototal, aes(x=celltype, y=count, fill=group)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", color="#4FE0DC", width=0.7) +
  scale_fill_manual(values=c("#4FE0DC", "white")) +
  theme_minimal()
ggsave(filename="oligototalhits.pdf", plot=oligototalbar, width=4.44, height=4.44, units="in")

```

#### plot gardener track of crispr pos region (region #18), one tailed p0.05, line at pvalue of 0.05
```{r}
#region 18
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="region18_onetailedp05_sigline.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides = FALSE, bg = NA)

pageGuideHide()


#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45779500, chromend = 45782500, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")



#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#plot CRISPR guides
plotRanges(data = beckycrispr, params = params_a, collapse = TRUE, y = 1, height = 0.25, boxHeight=0.25, fill = "#D88C9A")

#plot promoter capturec
plotRanges(data = maptcapturec_gluta, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#8E7DBE")

#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.2, height = .75, fill = "#86EAE6")

threshold <- -log10(0.05)
#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 2, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,3), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)

geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()
```

#### plot gardener track of crispr region 5 - one tailed p05, sig line
```{r}
#region 5
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="region5_onetailedp05_sigline.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45821000, chromend = 45823000, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")



#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#neeed to remove one crispr guide
#plot CRISPR guides
plotRanges(data = beckycrispr, params = params_a, collapse = TRUE, y = 1, height = 0.25, boxHeight=0.25, fill = "#D88C9A")

#plot promoter capturec
plotRanges(data = mapt_hicgluta, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#8E7DBE")

#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 2, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,4), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()
```

#### plot gardener region 9
```{r}
#region 9
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="region9_onetailedp05_sigline.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 46036535, chromend = 46039410, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")



#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#neeed to remove one crispr guide
#plot CRISPR guides
plotRanges(data = beckycrispr, params = params_a, collapse = TRUE, y = 1, height = 0.25, boxHeight=0.25, fill = "#D88C9A")

#plot promoter capturec
#plotRanges(data = mapt_hicgluta, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#8E7DBE")

#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold <- -log10(0.05)
#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 2, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,6), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()
```

#### plot gardener region 8
```{r}
#### plot gardener track of crispr region 5 - one tailed p05, sig line
#region 8
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="/cluster/home/rhauser/files/graphs/paperfigures/region8_onetailedp05_sigline.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 46029900, chromend = 46031100, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")



#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#neeed to remove one crispr guide
#plot CRISPR guides
plotRanges(data = beckycrispr, params = params_a, collapse = TRUE, y = 1, height = 0.25, boxHeight=0.25, fill = "#D88C9A")

#plot promoter capturec
plotRanges(data = maptcapturec_gluta, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#8E7DBE")

#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold <- -log10(0.05)

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 2, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,10), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()
```

####mapt prom region plotgardener
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="maptprom_onetailedp05_sigline.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45893536, chromend = 45895535, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#neeed to remove one crispr guide
#plot CRISPR guides
#plotRanges(data = beckycrispr, params = params_a, collapse = TRUE, y = 1, height = 0.25, boxHeight=0.25, fill = "#D88C9A")

#plot promoter capturec
#plotRanges(data = mapt_hicgluta, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#8E7DBE")

#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = cappedsignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,15), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45428787_45429857
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45428787_45429857.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45428787, chromend = 45429857, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = -0.3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = mapt_snmultiomics_brain, params = params_a, order.="random", y = -0.1, height = 0.75, fill = "#D88C9A")

#plot maptcapturec_gluta
plotRanges(data = maptcapturec_gluta, params = params_a, order. ="random", y = 0.1, height = .75, fill = "#8E7DBE")

plotRanges(data = maptluc, params = params_a, order. ="random", y = -0.5, height = .75, fill = "#037971")

#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 1.25, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,2), fontsize = 8, digits=3)

baclogplot <- plotSignal(data = bacmprasignal, params = params_a, negData = TRUE, y = 2.5, height = 0.75, fill = "#1c9c98", line = "#1c9c98")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=baclogplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = baclogplot,axisLine = TRUE, at=c(0,1.5), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 3.0, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()

```

####bri region 45848303_45849886
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45848303_45849886.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45848303, chromend = 45849886, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = mapt_snmultiomics_brain, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#D88C9A")

#plot luciferase
plotRanges(data = maptluc, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,4), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45873488_45874156
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45873488_45874156.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45873488, chromend = 45874156, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#plot luciferase
plotRanges(data = maptluc, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,3), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45940685_45941352
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45940685_45941352.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45940685, chromend = 45941352, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = mapt_snmultiomics_brain, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#D88C9A")

#plot luciferase
plotRanges(data = maptluc, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,2), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45942082_45942750
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45942082_45942750.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45942082, chromend = 45942750, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = mapt_snmultiomics_brain, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#D88C9A")

#plot luciferase
plotRanges(data = maptluc, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,4), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45971424_45972091
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45971424_45972091.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45971424, chromend = 45972091, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = maptcapturec_gaba, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#8E7DBE")

#plot luciferase
plotRanges(data = maptluc, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = avg_neurons, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,3.5), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45431444_45432656
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45431444_45432656.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45431444, chromend = 45432656, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = mapt_snmultiomics_brain, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#D88C9A")

#plot luciferase
plotRanges(data = maptluc, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = bacmprasignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,1.5), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45219148_45219936
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45219148_45219936.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45219148, chromend = 45219936, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = maptcapturec_gaba, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#8E7DBE")

#plot luciferase
#plotRanges(data = maptcapturec_gluta, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657
plotRanges(data = mapt_hicgluta, params = params_a, order.="random", y = 0.9, height = 0.75, fill = "#D88C9A")

#plot luciferase
plotRanges(data = maptcapturec_gluta, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#6852A3")

#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = bacmprasignal, params = params_a, negData = TRUE, y = 2.25, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,1.5), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45240986_45242338
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45240986_45242338.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45240986, chromend = 45242338, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
plotRanges(data = mapt_snmultiomics_culture, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#8E7DBE")

#plot luciferase
plotRanges(data = mapt_snmultiomics_brain, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#d88c9a")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = bacmprasignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,3), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45309210_45309877
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45309210_45309877.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45309210, chromend = 45309877, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")


 
#plot bricrispr
plotRanges(data = bricrisprpos_noprom, params = params_a, order. ="random", y = .3, height = .75, fill = "#0D3B66")

#neeed to remove one crispr guide
#plot mapt_snmultiomics_brain
#plotRanges(data = maptcapturec_gaba, params = params_a, order.="random", y = 0.5, height = 0.75, fill = "#8E7DBE")

#plot luciferase
plotRanges(data = maptluc, params = params_a, order. ="random", y = 0.7, height = .75, fill = "#037971")
#861657


#plot mpra data
#plotRanges(data = neurons_hits, params = params_a, order. ="random", y = 1.1, height = .75, fill = "#86EAE6")

threshold = -log10(0.05)

#cap signal
#cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))

#plot ompra
oligologplot <- plotSignal(data = bacmprasignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,2), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

#complex heatmap
```{r}

# Make sure log2FoldChange is numeric
filtered_data$log2FoldChange <- as.numeric(filtered_data$log2FoldChange)

# Pivot to wide format
mat <- dcast(filtered_data, gene ~ gRNA, value.var = "log2FoldChange", fill = 0)

gene_order <- c("DBF4B", "ADAM11", "GJC1", "EFTUD2", "GFAP", "KIF18B", "NMT1", "PLCD3",
                "ACBD4", "HEXIm1", "HEXIM2", "FMNL1", "MAP3K14", "ARHGAP27", 
                "PLEKHM1", "MAPT", "KANSL1")

# Desired gRNA (column) order
gRNA_order <- c("reg1", "reg2", "reg11", "reg14", "reg18", "reg4", "reg15", 
                "reg5", "reg13", "reg6", "reg7", "prom", "reg8", "reg9", "reg10")
# Convert to matrix, set rownames
mat_matrix <- as.matrix(mat[,-1])
rownames(mat_matrix) <- mat$gene
# Ensure that your gene and gRNA order values exist in the matrix
gene_order <- gene_order[gene_order %in% rownames(mat_matrix)]
gRNA_order <- gRNA_order[gRNA_order %in% colnames(mat_matrix)]

# Reorder the matrix
mat_matrix_ordered <- mat_matrix[gene_order, gRNA_order]

col_fun = colorRamp2(c(-2,-0.25,0,0.25,2), hcl_palette ="Purple-Yellow")

pdf("complexheatmap_200kb.pdf", width = 8, height = 8)
Heatmap(mat_matrix_ordered, col=col_fun, cluster_rows=FALSE, cluster_columns=FALSE, border=TRUE)
dev.off()

distmaptprom <- c(988706, 522197, 243399, 238653, 112735, 90799, 74247, 72162, 65587, 56834, 38368, 0, 136474, 143964, 178943)


column_col <- colorRamp2(c(0, 100000, 200000, 500000, 1000000), hcl_palette = "RdPu", reverse=TRUE)
column_annot <- HeatmapAnnotation(dist=distmaptprom, col=list(dist=column_col), border=TRUE)



pdf("complexheatmap_200kb.pdf", width = 8, height = 8)
Heatmap(mat_matrix_ordered, col=col_fun, cluster_rows=FALSE, cluster_columns=FALSE, border=TRUE, top_annotation=column_annot)
dev.off()

```



##Graph snvs overlap heks and neurons, default filt, implement a logFC absvalue of at least 0.1
```{r}
options(scipen=999)
library(scales)
snvsheks <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/heks/bcalm_snvs_labeled_heks_10bcthres.tsv")
snvsneurons <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/neurons/bcalm_snvs_labeled_neurons_10bcthres.tsv")

allsnvs <- full_join(
  snvsheks %>% rename_with(~paste0(.x, ".hek"), -element),
  snvsneurons %>% rename_with(~paste0(.x, ".neuron"), -element),
  by="element"
)

allsnvs <- allsnvs %>% mutate(sig = case_when(
  adj.P.Val.hek <=0.05 & adj.P.Val.neuron <= 0.05 & abs(logFC.hek) > 0.1 & abs(logFC.neuron) >0.1 ~ "sigboth",
  adj.P.Val.hek <=0.05 & abs(logFC.hek) >0.1 ~ "sighek",
  adj.P.Val.neuron <=0.05 & abs(logFC.neuron) >0.1 ~ "signeuron",
  TRUE ~ "ns"
))

write_tsv(allsnvs, "allsnvs.tsv")

allsnvs$sig <- factor(allsnvs$sig, 
                        levels = c("ns", "sighek", "signeuron", "sigboth"))

allsnvs <- allsnvs %>%
  arrange(sig)

snvscellcorr <- ggplot(allsnvs, aes(x = logFC.hek, y = logFC.neuron, color = sig)) +
  geom_point(size = 1.5) +
  scale_color_manual(name="Adj.P.Val & logFC cutoff",
    values = c(
      "sigboth"   = "#8332b5",
      "sighek"    = "#6acfdc",
      "signeuron" = "#ec5bb4",
      "ns"        = "#e0e0e0"
    ),
    labels = c(
    "ns"        = "None",
    "sighek"    = "HEKs",   # or your exact text
    "signeuron" = "Neurons",
    "sigboth"   = "HEKs & Neurons"
  )) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") +  
  geom_hline(yintercept = 0, linetype = "dotted", color = "black") +
  labs(
    x = "HEK293FT logFC",
    y = "Neuron logFC",
    title = "SNVs \nNeurons vs. HEK293FT"
  ) +
  theme_classic() +
  theme(legend.position = "right") +
  scale_y_continuous(limits=c(-1.2,0.8), breaks = seq(-1.2,0.8, by =0.2), labels=scales::label_number(accuracy=0.1)) +
  scale_x_continuous(limits=c(-0.8,0.8), breaks=seq(-0.8,0.8, by=0.2), labels=scales::label_number(accuracy=0.1))
ggsave(filename ="snvcorr_logFCcutoff_10bcthres.pdf", plot=snvscellcorr, width=5.44, height=4.44, units="in")

res <- cor.test(allsnvs$logFC.neuron, allsnvs$logFC.hek, method = "spearman", exact =FALSE)

res$estimate  # Spearmans's rho = 0.08645348 
res$p.value   # P-value 1.518072e-15

```

##Graph indels overlap heks and neurons, default filt, implement a logFC absvalue of at least 0.1
```{r}
indelsheks <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/heks/bcalm_indels_labeled_heks_10bcthres.tsv")
indelsneurons <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/neurons/bcalm_indels_labeled_neuronss_10bcthres.tsv")

allindels <- full_join(
  indelsheks %>% rename_with(~paste0(.x, ".hek"), -element),
  indelsneurons %>% rename_with(~paste0(.x, ".neuron"), -element),
  by="element"
)

allindels <- allindels %>% mutate(sig = case_when(
  adj.P.Val.hek <=0.05 & adj.P.Val.neuron <= 0.05 & abs(logFC.hek) >= 0.1 & abs(logFC.neuron) >=0.1 ~ "sigboth",
  adj.P.Val.hek <=0.05 & abs(logFC.hek) >=0.1 ~ "sighek",
  adj.P.Val.neuron <=0.05 & abs(logFC.neuron) >=0.1 ~ "signeuron",
  TRUE ~ "ns"
))


allindels$sig <- factor(allindels$sig, 
                        levels = c("ns", "sighek", "signeuron", "sigboth"))

allindels <- allindels %>%
  arrange(sig)

indelcellcorr <- ggplot(allindels, aes(x = logFC.hek, y = logFC.neuron, color = sig)) +
  geom_point(size = 1.5) +
  scale_color_manual(name="Adj.P.Val & logFC cutoff",
    values = c(
      "sigboth"   = "#8332b5",
      "sighek"    = "#6acfdc",
      "signeuron" = "#ec5bb4",
      "ns"        = "#e0e0e0"
    ),
    labels = c(
    "ns"        = "None",
    "sighek"    = "HEKs",   # or your exact text
    "signeuron" = "Neurons",
    "sigboth"   = "HEKs & Neurons"
  )) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") +  
  geom_hline(yintercept = 0, linetype = "dotted", color = "black") +
  labs(
    x = "HEK293FT logFC",
    y = "Neuron logFC",
    title = "InDels \nNeurons vs. HEK293FT"
  ) +
  ylim(-0.6,0.4) +
  theme_classic() +
  theme(legend.position = "right")
ggsave(filename="indelcorr_logFCcutoff_10bcthres.pdf", plot=indelcellcorr, width=5.44, height=4.44, units="in")


```

##neuron mapt promoter heatmap
```{r}
mapt <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/neurons/bcalm_mapt_neurons_10bcthres.tsv")

mapt <- mapt %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < 0 ~ "sig_down",
      adj.P.Val > 0.05 ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

mapt <- mapt %>% mutate(position = as.numeric(str_extract(element, "(?<=maptprom_chr17_)[0-9]+")))

bcalmmapt <- ggplot(mapt, aes(x = position, y = allele, fill = sig2)) +
  geom_tile() +
  theme_minimal() +  # Use a minimal theme
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    ) +  # Rotate x-axis labels for readability
    labs(,
      fill = "active control v variant"
    ) +
  scale_fill_manual(values = c(
    "ns" = "gray85" ,
    "sig_up" = "tomato",
    "sig_down" ="dodgerblue"),
    na.value = "white"
  )
ggsave(filename="maptheatmap_bcalm_10bcthres.pdf", plot=bcalmmapt, width=30, height=4.44, units="in")




#implementing a logFC thres of 0.1
mapt <- mapt %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0.1 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < -0.1 ~ "sig_down",
      adj.P.Val > 0.05 ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))
bcalmmapt <- ggplot(mapt, aes(x = position, y = allele, fill = sig2, text=position)) +
  geom_tile() +
  theme_minimal() +  # Use a minimal theme
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    ) +  # Rotate x-axis labels for readability
    labs(,
      fill = "active control v variant"
    ) +
  scale_fill_manual(values = c(
    "ns" = "gray85" ,
    "sig_up" = "tomato",
    "sig_down" ="dodgerblue"),
    na.value = "white"
  )
ggsave(filename="maptheatmap_bcalm_10bcthres_logFCcutoff.pdf", plot=bcalmmapt, width=30, height=4.44, units="in")

library(plotly)
bcalmmaptp <- ggplotly(bcalmmapt, tooltip="text")

htmlwidgets::saveWidget(bcalmmaptp, file = "maptheatmap_bcalm_10bcthres_logFCcutoff.html", selfcontained = TRUE)



```

##complex heatmap logfc
```{r}
# Make sure log2FoldChange is numeric
mapt$logFC <- as.numeric(mapt$logFC)

# Pivot to wide format
maptwide <- dcast(mapt, allele ~ position, value.var = "logFC", fill = 0)

posorder <- mapt %>% pull(position) %>% sort()
posorder <- unique(posorder)

# Desired row order
alleleorder <- c("E", "B", "C", "D")

# Convert to matrix, set rownames
mat_matrix <- as.matrix(maptwide[,-1])
rownames(mat_matrix) <- maptwide$allele
# Ensure that your allele and position order values exist in the matrix
# Ensure that your allele and position order values exist in the matrix
allele_order <- alleleorder[alleleorder %in% rownames(mat_matrix)]
pos_order <- posorder[posorder %in% colnames(mat_matrix)]
pos_order <- as.character(pos_order)



# Reorder the matrix
mat_matrix_ordered <- mat_matrix[allele_order, pos_order]

col_fun = colorRamp2(c(-0.5,0,0.5), c("dodgerblue", "white", "tomato"))


pdf("complexheatmap_maptvarlogfc.pdf", width = 30, height = 4.44)
Heatmap(mat_matrix_ordered, col=col_fun, cluster_rows=FALSE, cluster_columns=FALSE, border=TRUE)
dev.off()






```

##complex heatmap heks maptprom
```{r}
library(ComplexHeatmap)
library(reshape2)
maptheks <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/heks/bcalm_mapt_heks_10bcsthres.tsv")
maptheks <- maptheks %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0.1 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < -0.1 ~ "sig_down",
      adj.P.Val > 0.05 ~ "ns",
      adj.P.Val <= 0.05 & logFC >= -0.1 & logFC <= 0.1 ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

maptheks <- maptheks %>% mutate(position = as.numeric(str_extract(element, "(?<=maptprom_chr17_)[0-9]+")))


maptheks$sig2 <- as.character(maptheks$sig2)

# Pivot to wide format on sig2
mapthekswide <- dcast(maptheks, allele ~ position, value.var = "sig2", fill = NA)

posorder    <- sort(unique(maptheks$position))
alleleorder <- c("E", "B", "C", "D")

# Convert to matrix and set rownames
mat_matrix <- as.matrix(mapthekswide[ , -1])
rownames(mat_matrix) <- as.character(mapthekswide$allele)

# Ensure rows/cols exist
allele_order <- alleleorder[alleleorder %in% rownames(mat_matrix)]
pos_order    <- as.character(posorder[posorder %in% colnames(mat_matrix)])

# Reorder matrix
mat_matrix_ordered <- mat_matrix[allele_order, pos_order]

# Define discrete color mapping
category_colors <- c(
  sig_up   = "tomato",
  sig_down = "dodgerblue",
  ns       = "gray90"
)

# Plot heatmap (with NA = white)
pdf("complexheatmap_maptvar_sig_heks2.pdf", width = 30, height = 4.44)
Heatmap(
  mat_matrix_ordered,
  col     = category_colors,
  na_col  = "white",
  cluster_rows    = FALSE,
  cluster_columns = FALSE,
  border = TRUE
)
dev.off()


```


##mapt promoter common between cell types
```{r}
maptheks_renamed <- maptheks %>%
  rename_with(~ paste0(.x, ".heks"), .cols = -c(position, allele))

maptcombined <- mapt %>%
  inner_join(maptheks_renamed, by = c("position", "allele"))

maptcombined <- maptcombined %>% mutate(
    shared = case_when(
      sig2 == "sig_up" & sig2.heks == "sig_up" ~ "sig_up",
      sig2 == "sig_down" & sig2.heks == "sig_down" ~ "sig_down",
      sig2 == "sig_down" & sig2.heks == "sig_up" ~ "ns",
      sig2 == "sig_up" & sig2.heks == "sig_down" ~ "ns",
      sig2 == "sig_up" & sig2.heks == "ns" ~ "ns",
      sig2 == "sig_down" & sig2.heks == "ns" ~ "ns",
      sig2 == "ns" & sig2.heks == "sig_up" ~ "ns",
      sig2 == "ns" & sig2.heks == "sig_down" ~ "ns",
      sig2 == "ns" & sig2.heks == "ns" ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

maptcombined$shared <- as.character(maptcombined$shared)

# Pivot to wide format on sig2
maptcombinedwide <- dcast(maptcombined, allele ~ position, value.var = "shared", fill = NA)

posorder    <- sort(unique(maptcombined$position))
alleleorder <- c("E", "B", "C", "D")

# Convert to matrix and set rownames
mat_matrix <- as.matrix(maptcombinedwide[ , -1])
rownames(mat_matrix) <- as.character(maptcombinedwide$allele)

# Ensure rows/cols exist
allele_order <- alleleorder[alleleorder %in% rownames(mat_matrix)]
pos_order    <- as.character(posorder[posorder %in% colnames(mat_matrix)])

# Reorder matrix
mat_matrix_ordered <- mat_matrix[allele_order, pos_order]

# Define discrete color mapping
category_colors <- c(
  sig_up   = "tomato",
  sig_down = "dodgerblue",
  ns       = "gray90"
)

# Plot heatmap (with NA = white)
pdf("complexheatmap_maptvar_shared_heksandneurons.pdf", width = 30, height = 4.44)
Heatmap(
  mat_matrix_ordered,
  col     = category_colors,
  na_col  = "white",
  cluster_rows    = FALSE,
  cluster_columns = FALSE,
  border = TRUE
)
dev.off()


```



##mapt promoter neuron specific
```{r}
maptheks_renamed <- maptheks %>%
  rename_with(~ paste0(.x, ".heks"), .cols = -c(position, allele))

maptcombined <- mapt %>%
  inner_join(maptheks_renamed, by = c("position", "allele"))

maptcombined <- maptcombined %>% mutate(
    neuronspecific = case_when(
      sig2 == "sig_up" & sig2.heks == "sig_up" ~ "ns",
      sig2 == "sig_down" & sig2.heks == "sig_down" ~ "ns",
      sig2 == "sig_down" & sig2.heks == "sig_up" ~ "sig_down",
      sig2 == "sig_up" & sig2.heks == "sig_down" ~ "sig_up",
      sig2 == "sig_up" & sig2.heks == "ns" ~ "sig_up",
      sig2 == "sig_down" & sig2.heks == "ns" ~ "sig_down",
      sig2 == "ns" & sig2.heks == "sig_up" ~ "ns",
      sig2 == "ns" & sig2.heks == "sig_down" ~ "ns",
      sig2 == "ns" & sig2.heks == "ns" ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

maptcombined$neuronspecific <- as.character(maptcombined$neuronspecific)

# Pivot to wide format on sig2
maptcombinedwide <- dcast(maptcombined, allele ~ position, value.var = "neuronspecific", fill = NA)

posorder    <- sort(unique(maptcombined$position))
alleleorder <- c("E", "B", "C", "D")

# Convert to matrix and set rownames
mat_matrix <- as.matrix(maptcombinedwide[ , -1])
rownames(mat_matrix) <- as.character(maptcombinedwide$allele)

# Ensure rows/cols exist
allele_order <- alleleorder[alleleorder %in% rownames(mat_matrix)]
pos_order    <- as.character(posorder[posorder %in% colnames(mat_matrix)])

# Reorder matrix
mat_matrix_ordered <- mat_matrix[allele_order, pos_order]

# Define discrete color mapping
category_colors <- c(
  sig_up   = "tomato",
  sig_down = "dodgerblue",
  ns       = "gray90"
)

# Plot heatmap (with NA = white)
pdf("complexheatmap_maptvar_neuronspecific.pdf", width = 30, height = 4.44)
Heatmap(
  mat_matrix_ordered,
  col     = category_colors,
  na_col  = "white",
  cluster_rows    = FALSE,
  cluster_columns = FALSE,
  border = TRUE
)
dev.off()


```

#mapt logfc by base, gain of activity region
```{r}
#label base ids
maptkey <- read_tsv("/cluster/home/rhauser/files/rmhex03/design/fiveallelesmpradesign.tsv")
mapt <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/neurons/bcalm_mapt_neurons_10bcthres.tsv")

mapt <- mapt %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < 0 ~ "sig_down",
      adj.P.Val > 0.05 ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

mapt <- mapt %>% mutate(position = as.numeric(str_extract(element, "(?<=maptprom_chr17_)[0-9]+")))

maptkey <- maptkey %>% filter(str_detect(id, "mapt"))
maptkey <- maptkey %>% filter(!str_detect(id, "ref"))
maptkey <- maptkey %>% filter(posControl != TRUE)
maptkey <- maptkey %>% filter(negControl != TRUE)
maptkeylong <- maptkey %>% pivot_longer(cols = starts_with("allele"),
               names_to = "allele",
               values_to = "present") %>%
  # keep only rows where that allele is TRUE
  filter(present) %>%
  # clean allele column: remove "allele" prefix to just leave A, B, C...
  mutate(allele = sub("allele", "", allele)) %>%
  select(-present)
maptkey_long <- maptkey %>%
  pivot_longer(cols = starts_with("allele"),
  names_to = "allele",
  values_to = "has_allele") %>%
  filter(has_allele == TRUE) %>%
  mutate(allele = sub("allele", "", allele))

mapt_with_id <- mapt %>%
  left_join(
    maptkey_long %>% select(id, element, allele),
    by = c("element", "allele")
  )
#mapt_with_id <- mapt_with_id %>% filter(allele != "E")
mapt_with_id <- mapt_with_id %>% separate(id,
           into = c("prefix", "chr", "pos", "altseq", "base_ref", "to", "base_alt"),
           sep = "_",
           remove = FALSE,
           extra = "merge") %>%
  select(-prefix, -chr, -pos, -altseq, -to)
mapt_with_id <- mapt_with_id %>% mutate(base_alt = replace(base_alt, is.na(base_alt), "deletion"))
mapt_with_id_filt30bp <- mapt_with_id %>% filter(between(position, 45894673, 45894702))
mapt_with_id_filt30bp <- mapt_with_id_filt30bp %>% arrange(by=adj.P.Val)
mapt_with_id_filt30bp <- mapt_with_id_filt30bp %>% mutate(sig = case_when(
  adj.P.Val <=0.05 ~ "sig alt",
  TRUE ~ "ns"
))
mapt_with_id_filt30bp <- mapt_with_id_filt30bp %>%
  dplyr::mutate(
    base_alt = factor(base_alt, levels = c("A", "C", "G", "T", "deletion"))
  )
mapt_fc30bp <- ggplot(mapt_with_id_filt30bp, aes(fill=base_alt, y=logFC, x=position, alpha=sig)) +
  geom_bar(position="dodge", stat="identity", width=0.7) +
  theme_minimal() +
  scale_fill_manual(
    values = c(
      "A" = "#1b998b", 
      "T" = "#ff6347",  
      "C" = "#5cadff", 
      "G" = "#fecb34",
      "deletion" = "gray55"
    )) +
  scale_alpha_manual(values=c("ns" = 0.25, "sig alt" = 1)) +
  geom_hline(yintercept=0)
ggsave(filename="maptbcalm_logfc_30bases_gainactivity.pdf", plot=mapt_fc30bp, width=15, height=4.44, units="in")


mapt_fc30bpogcolors <- ggplot(mapt_with_id_filt30bp, aes(fill=base_alt, y=logFC, x=position, alpha=sig)) +
  geom_bar(position="dodge", stat="identity", width=0.7) +
  theme_minimal() +
  scale_fill_manual(
    values = c(
      "A" = "#00811b", 
      "T" = "#d00001",  
      "C" = "#2000c7", 
      "G" = "#ffb32c",
      "deletion" = "gray75"
    )) +
  scale_alpha_manual(values=c("ns" = 0.25, "sig alt" = 1)) +
  geom_hline(yintercept=0)
ggsave(filename="maptbcalm_logfc_30bases_gainactivityOGcolors.pdf", plot=mapt_fc30bpogcolors, width=15, height=4.44, units="in")
```


##mapt logfc by base, loss of activity region
```{r}
#label base ids
maptkey <- read_tsv("/cluster/home/rhauser/files/rmhex03/design/fiveallelesmpradesign.tsv")


maptkey <- maptkey %>% filter(str_detect(id, "mapt"))
maptkey <- maptkey %>% filter(!str_detect(id, "ref"))
maptkey <- maptkey %>% filter(posControl != TRUE)
maptkey <- maptkey %>% filter(negControl != TRUE)
maptkeylong <- maptkey %>% pivot_longer(cols = starts_with("allele"),
               names_to = "allele",
               values_to = "present") %>%
  # keep only rows where that allele is TRUE
  filter(present) %>%
  # clean allele column: remove "allele" prefix to just leave A, B, C...
  mutate(allele = sub("allele", "", allele)) %>%
  select(-present)
maptkey_long <- maptkey %>%
  pivot_longer(cols = starts_with("allele"),
  names_to = "allele",
  values_to = "has_allele") %>%
  filter(has_allele == TRUE) %>%
  mutate(allele = sub("allele", "", allele))

mapt_with_id <- mapt %>%
  left_join(
    maptkey_long %>% select(id, element, allele),
    by = c("element", "allele")
  )
#mapt_with_id <- mapt_with_id %>% filter(allele != "E")
mapt_with_id <- mapt_with_id %>% separate(id,
           into = c("prefix", "chr", "pos", "altseq", "base_ref", "to", "base_alt"),
           sep = "_",
           remove = FALSE,
           extra = "merge") %>%
  select(-prefix, -chr, -pos, -altseq, -to)
mapt_with_id <- mapt_with_id %>% mutate(base_alt = replace(base_alt, is.na(base_alt), "deletion"))
mapt_with_id_filt30bp <- mapt_with_id %>% filter(between(position, 45894270, 45894299))
mapt_with_id_filt30bp <- mapt_with_id_filt30bp %>% arrange(by=adj.P.Val)
mapt_with_id_filt30bp <- mapt_with_id_filt30bp %>% mutate(sig = case_when(
  adj.P.Val <=0.05 ~ "sig alt",
  TRUE ~ "ns"
))
mapt_with_id_filt30bp <- mapt_with_id_filt30bp %>%
  dplyr::mutate(
    base_alt = factor(base_alt, levels = c("A", "C", "G", "T", "deletion"))
  )
mapt_fc30bp <- ggplot(mapt_with_id_filt30bp, aes(fill=base_alt, y=logFC, x=position, alpha=sig)) +
  geom_bar(position="dodge", stat="identity", width=0.7) +
  theme_minimal() +
  scale_fill_manual(
    values = c(
      "A" = "#1b998b", 
      "T" = "#ff6347",  
      "C" = "#5cadff", 
      "G" = "#fecb34",
      "deletion" = "gray55"
    )) +
  scale_alpha_manual(values=c("ns" = 0.25, "sig alt" = 1)) +
  geom_hline(yintercept=0)
ggsave(filename="maptbcalm_logfc_30bases.pdf", plot=mapt_fc30bp, width=15, height=4.44, units="in")

mapt_fc30bpogcolors <- ggplot(mapt_with_id_filt30bp, aes(fill=base_alt, y=logFC, x=position, alpha=sig)) +
  geom_bar(position="dodge", stat="identity", width=0.7) +
  theme_minimal() +
  scale_fill_manual(
    values = c(
      "A" = "#00811b", 
      "T" = "#d00001",  
      "C" = "#2000c7", 
      "G" = "#ffb32c",
      "deletion" = "gray75"
    )) +
  scale_alpha_manual(values=c("ns" = 0.25, "sig alt" = 1)) +
  geom_hline(yintercept=0)
ggsave(filename="maptbcalm_logfc_30bases_lossactivityOGcolors.pdf", plot=mapt_fc30bpogcolors, width=15, height=4.44, units="in")


```

#showing logfc for the dot plots
```{r}

bricrispr_45971424_45972091 <- ggplot(snvsindels, aes(x=position, y = logFC, color = sig2 )) +
  geom_point(size=3) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85" ,
    "sig_up" = "tomato",
    "sig_down" ="dodgerblue"),
    na.value = "white"
  ) +
  coord_cartesian(xlim = c(45971424, 45972091), ylim=c(-0.5,0.5))
ggsave(filename="bricrispr_45971424_45972091_logfc.pdf", plot=bricrispr_45971424_45972091, width=10, height=2.22, units="in")



snvsindels <- snvsindels %>% mutate(
    snpID = str_replace(
      element,
      ".*?(chr[0-9XYM]+)_([0-9]+)_(?:altseq_)?([A-Za-z]+)_to_([A-Za-z]+)$",
      "\\1:\\2:\\3:\\4"
    )
  )
snvsindels <- snvsindels %>%
  mutate(sig2 = factor(sig2, levels = c("ns", "sig_down", "sig_up"))) %>%
  arrange(sig2)

bricrispr_45971424_45972091 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45971424, 45972091), ylim = c(-0.25, 0.25)) +
  scale_x_continuous(breaks=c(45971424, 45972091), labels=c("45971424", "45972091"))
ggsave(filename="bricrispr_45971424_45972091_logfc.pdf", plot=bricrispr_45971424_45972091, width=10, height=2.22, units="in")

bricrispr_45942082_45942750 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45942082, 45942750), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(45942082, 45942750), labels=c("45942082", "45942750"))
ggsave(filename="bricrispr_45942082_45942750_logfc.pdf", plot=bricrispr_45942082_45942750, width=10, height=2.22, units="in")


bricrispr_45940685_45941352 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45940685, 45941352), ylim=c(-0.3,0.3)) +
  scale_x_continuous(breaks=c(45940685, 45941352), labels=c("45940685", "45941352"))
ggsave(filename="bricrispr_45940685_45941352_logfc.pdf", plot=bricrispr_45940685_45941352, width=10, height=2.22, units="in")

bricrispr_45873488_45874156 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45873488, 45874156), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(45873488, 45874156), labels=c("45873488", "45874156"))
ggsave(filename="bricrispr_45873488_45874156_logfc.pdf", plot=bricrispr_45873488_45874156, width=10, height=2.22, units="in")


bricrispr_45848303_45849886 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45848303, 45849886), ylim=c(-0.25,0.25)) +
  scale_x_continuous(breaks=c(45848303, 45849886), labels=c("45848303", "45849886"))
ggsave(filename="bricrispr_45848303_45849886_logfc.pdf", plot=bricrispr_45848303_45849886, width=10, height=2.22, units="in")

bricrispr_45431444_45432656 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45431444, 45432656), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(45431444, 45432656), labels=c("45431444", "45432656"))
ggsave(filename="bricrispr_45431444_45432656_logfc.pdf", plot=bricrispr_45431444_45432656, width=10, height=2.22, units="in")

bricrispr_45428787_45429857 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45428787, 45429857), ylim=c(-0.3,0.3)) +
    scale_x_continuous(breaks=c(45428787, 45429857), labels=c("45428787", "45429857"))
ggsave(filename="bricrispr_45428787_45429857_logfc.pdf", plot=bricrispr_45428787_45429857, width=10, height=2.22, units="in")

bricrispr_45309210_45309877 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45309210, 45309877), ylim=c(-0.25,0.25)) +
  scale_x_continuous(breaks=c(45309210, 45309877), labels=c("45309210", "45309877"))
ggsave(filename="bricrispr_45309210_45309877_logfc.pdf", plot=bricrispr_45309210_45309877, width=10, height=2.22, units="in")

bricrispr_45240986_45242338 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45240986, 45242338), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(45240986, 45242338), labels=c("45240986", "45242338"))
ggsave(filename="bricrispr_45240986_45242338_logfc.pdf", plot=bricrispr_45240986_45242338, width=10, height=2.22, units="in")

bricrispr_45219148_45219936 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45219148, 45219936), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(45219148, 45219936), labels=c("45219148", "45219936"))
ggsave(filename="bricrispr_45219148_45219936_logfc.pdf", plot=bricrispr_45219148_45219936, width=10, height=2.22, units="in")


region5 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45821000, 45823000), ylim=c(-0.25,0.25)) +
  scale_x_continuous(breaks=c(45821000, 45823000), labels=c("45821000", "45823000"))
ggsave(filename="crispr_region5_logfc.pdf", plot=region5, width=10, height=2.22, units="in")

region18 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45780724, 45781806), ylim=c(-0.25,0.25)) +
  scale_x_continuous(breaks=c(45780724, 45781806), labels=c("45780724", "45781806"))
ggsave(filename="crispr_region18_logfc.pdf", plot=region18, width=10, height=2.22, units="in")

region9 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(46037318, 46038610), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(46037318, 46038610), labels=c("46037318", "46038610"))
ggsave(filename="crispr_region9_logfc.pdf", plot=region9, width=10, height=2.22, units="in")


region8 <- ggplot(snvsindels, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(46029900, 46031100), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(46029900, 46031100), labels=c("46029900", "46031100"))
ggsave(filename="crispr_region8_logfc.pdf", plot=region8, width=10, height=2.22, units="in")

#add in MAPT SNVs in ADSP
variantcounts <- read_tsv("/cluster/home/rhauser/files/rmhex03/variantinfo/variantinfo.tsv")
variantcounts <- variantcounts %>% rename("snpID" = "SNP")
variantcounts <- variantcounts %>%
  mutate(snpID = paste0("chr", snpID))
maptlabeled <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/neurons/bcalm_mapt_neurons_10bcthres_withlabels.tsv")
#remove 5bp deletion
maptlabeled <- maptlabeled %>% filter(allele != "E")
maptlabeled <- maptlabeled %>% mutate(
    snpID = str_replace(
      id,
      ".*?(chr[0-9XYM]+)_([0-9]+)_altseq[0-9]*_([A-Za-z]+)_to_([A-Za-z]+)$",
      "\\1:\\2:\\3:\\4"
    )
  )


ADSPmaptpromsnvs <- maptlabeled %>% inner_join(variantcounts, by="snpID")
ADSPmaptpromsnvs_labels <- ADSPmaptpromsnvs %>% select(snpID, CHR, pos, base_ref, base_alt)
ADSPmaptpromsnvs_labels <- ADSPmaptpromsnvs_labels %>% arrange(by=pos)
write_tsv(ADSPmaptpromsnvs_labels, "MAPTpromoterADSPvariants.tsv")

ADSPmaptpromsnvs <- ADSPmaptpromsnvs %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < 0 ~ "sig_down",
      #adj.P.Val > 0.05 ~ "ns",
      TRUE    ~ "ns"  # Fallback if none are TRUE
    ))
write_tsv(ADSPmaptpromsnvs, "ADSPmaptpromsnvs.tsv")


ADSPmaptpromsnvs <- ADSPmaptpromsnvs %>% rename("position" = "pos")
ADSPmaptpromsnvs <- ADSPmaptpromsnvs %>% select(element, logFC, AveExpr, t, P.Value, adj.P.Val, B, position, sig2, snpID)
snvsindels <- snvsindels %>% select(!group)
snvsindels2 <- rbind(snvsindels, ADSPmaptpromsnvs)

snvsindels2 <- snvsindels2 %>% mutate(sig2 = factor(sig2, levels = c("ns", "sig_down", "sig_up"))) %>%
  arrange(sig2)


maptindels <- ggplot(snvsindels2, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  geom_text(
    aes(label = ifelse(sig2 %in% c("sig_up", "sig_down"), snpID, "")),
    vjust = -1,
    size = 3,
    show.legend = FALSE
  ) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45893536, 45895535), ylim=c(-0.3,0.3)) +
  scale_x_continuous(breaks=c(45893536, 45895535), labels=c("45893536", "45895535"))
ggsave(filename="mapt_ADSPvars.pdf", plot=maptindels, width=10, height=2.22, units="in")



#filter out AC of 0
variantcounts <- variantcounts %>%
  filter(AD_minorCount != 0 & CTL_minorCount != 0)

ADSPmaptpromsnvs <- maptlabeled %>% inner_join(variantcounts, by="snpID")
ADSPmaptpromsnvs_labels <- ADSPmaptpromsnvs %>% select(snpID, CHR, pos, base_ref, base_alt)
ADSPmaptpromsnvs_labels <- ADSPmaptpromsnvs_labels %>% arrange(by=pos)
write_tsv(ADSPmaptpromsnvs_labels, "MAPTpromoterADSPvariants.tsv")

ADSPmaptpromsnvs <- ADSPmaptpromsnvs %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < 0 ~ "sig_down",
      #adj.P.Val > 0.05 ~ "ns",
      TRUE    ~ "ns"  # Fallback if none are TRUE
    ))
write_tsv(ADSPmaptpromsnvs, "ADSPmaptpromsnvs.tsv")


ADSPmaptpromsnvs <- ADSPmaptpromsnvs %>% rename("position" = "pos")
ADSPmaptpromsnvs <- ADSPmaptpromsnvs %>% select(element, logFC, AveExpr, t, P.Value, adj.P.Val, B, position, sig2)
#snvsindels <- snvsindels %>% select(!group)
snvsindels2 <- rbind(snvsindels, ADSPmaptpromsnvs)

snvsindels2 <- snvsindels2 %>% mutate(sig2 = factor(sig2, levels = c("ns", "sig_down", "sig_up"))) %>%
  arrange(sig2)


maptindels <- ggplot(snvsindels2, aes(x = position, y = logFC, color = sig2)) +
  geom_point(size = 3) +
  theme_minimal() +
  scale_color_manual(values = c(
    "ns" = "gray85",
    "sig_up" = "tomato",
    "sig_down" = "dodgerblue"
  ),
  na.value = "white") +
  coord_cartesian(xlim = c(45893536, 45895535), ylim=c(-0.3,0.3)) +
  scale_x_continuous(breaks=c(45893536, 45895535), labels=c("45893536", "45895535"))
ggsave(filename="mapt_ADSPvars_noAC0.pdf", plot=maptindels, width=10, height=2.22, units="in")


```

#BrainTF
```{r}
library(GenomicRanges)
library(ggrepel)
library(rtracklayer)

#read in data
union_files<-as.list(readLines("/cluster/home/aanderson/myers/BrainTF/path_to_unlist_csv.txt"))
m=gsub(".csv","", unlist(union_files))
m=sapply(strsplit(m, "/"),`[`,8)

peaks<-c()
union_list=lapply(union_files,function(x){
  tmp<-GRanges(read.csv(x))
 # peaks<-c(peaks, length(tmp[which(tmp$name=="NFIC"),]))
})

peaks<-do.call(c, union_list)
peaks<-peaks[order(-width(peaks)),]

Pileup_wName <- function(object, region, highlight_names = NULL) {
  peaks_df <- as.data.frame(subsetByOverlaps(object, region))
  peaks_df$tag <- paste0(peaks_df$name, "-", peaks_df$type)
  peaks_df <- peaks_df[!duplicated(peaks_df$tag), ]
  
  # Clip partially overlapping peaks so that they fit within the region
  peaks_df$start <- pmax(peaks_df$start, start(region))
  peaks_df$end   <- pmin(peaks_df$end, end(region))

  # Set IRanges for overlap staggering
  peak_ranges <- IRanges(start = peaks_df$start, end = peaks_df$end)
  overlap_levels <- disjointBins(peak_ranges)
  peaks_df$y_level <- overlap_levels

  # Label certain names if requested
  peaks_df$label <- if (!is.null(highlight_names)) {
    ifelse(peaks_df$name %in% highlight_names, peaks_df$name, "")
  } else {
    ""
  }

  # Build plot
  p <- ggplot(peaks_df) +
    geom_segment(aes(x = start, xend = end, y = y_level, yend = y_level, color = type), size = 3) +
    geom_text_repel(
      data = subset(peaks_df, label != ""),
      aes(x = (start + end) / 2, y = y_level, label = label),
      size = 3,
      box.padding = 0.3,
      point.padding = 0.2,
      segment.color = "grey50",
      max.overlaps = 100
    ) +
    theme_classic()+
    scale_y_continuous(name = "Overlap Level") +
    xlab(paste0(as.character(seqnames(region)), " position (bp)")) +
    xlim(c(start(region), end(region))) +
    geom_vline(xintercept = c(45893336, 45895735))  +
    ylim(c(0, 40)) +
    theme(legend.position = "none")

  return(p)
}

region2<-GRanges("chr17:45893336-45895735")
p1 <- Pileup_wName(peaks[which(peaks$type=="NEUN"),], region2)
ggsave("MAPTpromoter_brainTF_neun.pdf", plot=p1, height=2.22, width=10)


peaks

```

#boxplots of vars of interest
```{r}
neuroncounts <- read_tsv("/cluster/home/rhauser/MPRAsnakeflow/results/experiments/rmhex03Count/reporter_experiment.barcode.Neurons.fromFile.default.min_oligo_threshold_10.tsv")

merged_neuroncounts <- neuroncounts %>%
  select(-barcode) %>%
  group_by(oligo_name) %>%
  summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) %>%
  ungroup()

merged_neuroncounts_filt <- merged_neuroncounts %>%
  filter(
    dna_count_1 >= 50,
    dna_count_2 >= 50,
    dna_count_3 >= 50,
    dna_count_4 >= 50
  )

#ratios for RNA:DNA ratio calc neurons
#rep 1 lines in RNA: 4518933728, lines in DNA: 2091153840, 2.160976224 multiplier
#rep 2 lines in RNA: 4603339708, lines in DNA: 1907240732, 2.41361231 multiplier 
#rep 3 lines in RNA: 4452619052, lines in DNA: 1596127992, 2.789637845 multiplier
#rep 4 lines in RNA: 4839694712, lines in DNA: 1698000164, 2.850232182 multiplier

merged_neuroncounts_rnadna <- merged_neuroncounts %>% mutate(rnadna1 = ((rna_count_1/2.160976224)/dna_count_1),
                                                             rnadna2 = ((rna_count_2/2.41361231)/dna_count_2),
                                                             rnadna3 = ((rna_count_3/2.789637845)/dna_count_3),
                                                             rnadna4 = ((rna_count_3/2.850232182)/dna_count_4))

merged_neuroncounts_rnadna <- merged_neuroncounts_rnadna %>%
  mutate(mean_rnadna = rowMeans(across(c(rnadna1, rnadna2, rnadna3, rnadna4)), na.rm = TRUE))

merged_neuroncounts_rnadna <- merged_neuroncounts_rnadna %>% mutate(log2rnadna = log2(mean_rnadna))
merged_neuroncounts_rnadna <- merged_neuroncounts_rnadna %>% rename("label" = "oligo_name")

chr17_45894776 <- merged_neuroncounts_rnadna %>% filter(str_detect(label, "45894776"))
chr17_45894776 <- chr17_45894776 %>%
  # filter rows where “45894776” appears in label
  filter(str_detect(label, "45894776")) %>%
  # add ‘group’ column
  mutate(
    group = case_when(
      str_detect(label, "_altseq\\d+_([A-Z])_to_([A-Z])$") ~ str_extract(label, "_altseq\\d+_([A-Z])_to_([A-Z])$") %>%
                                                 str_remove("_altseq\\d+_") %>%
                                                 str_replace("_to_", ":"),
      str_detect(label, "_deletion$")       ~ "deletion",
      str_detect(label, "_referenceseq$")   ~ "reference",
      TRUE                                   ~ NA_character_
    )
  )

chr17_45894776 <- chr17_45894776 %>%
  pivot_longer(
    cols = c(rnadna1, rnadna2, rnadna3, rnadna4),
    names_to = "rnadna_source",     # optional: to keep track where it came from
    values_to = "rnadna"            # this will be your single column of values
  )


chr17_45894776 <- chr17_45894776 %>% mutate(group = factor(group, levels = c("reference", "G:A", "G:C", "G:T", "deletion"))) %>%
  arrange(group)


chr17_45894776plot <- ggplot(chr17_45894776, aes(x=group, y=rnadna)) +
  theme_classic() +
stat_boxplot(geom = "errorbar", width = 0.3, lwd=1.2, color=c("#d3d3d3", "#ca57ff", "#ff78c6", "#ff78c6", "#ff78c6")) +
    geom_boxplot(fill=c("#e0e0e0", "#d270ff", "#ff91d1", "#ff91d1", "#ff91d1"), color=c("#d3d3d3", "#ca57ff", "#ff78c6", "#ff78c6", "#ff78c6"), lwd=1.2) +
stat_summary(
    fun = median, 
    geom = "crossbar", 
    width = 0.7, 
    color = "white", 
    linewidth = 0.7
  )

ggsave(filename="rs11575896_rnadna.pdf", plot=chr17_45894776plot , width=4, height=4, units="in")



```

#alligning alphagenome, promoterai, and mpradata 2
```{r}
promai <- read_tsv("/cluster/home/rhauser/files/rmhex03/AIfromHenry/promoterAI_promoter_mutagenesis_results.tsv")
alphagenomeprom <- read_csv("/cluster/home/rhauser/files/rmhex03/AIfromHenry/alphagenome_mapt_promoter_mutagenesis_glut.csv")
overlapup <- read_csv("/cluster/home/rhauser/files/rmhex03/AIfromHenry/maptprom_snvs_up_overlap.csv")
overlapdown <- read_csv("/cluster/home/rhauser/files/rmhex03/AIfromHenry/maptprom_snvs_down_overlap.csv")
maptlabeled <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/neurons/bcalm_mapt_neurons_10bcthres_withlabels.tsv")
#remove 5bp deletion
maptlabeled <- maptlabeled %>% filter(allele != "E")
maptlabeled <- maptlabeled %>% mutate(
    snpID = str_replace(
      id,
      ".*?(chr[0-9XYM]+)_([0-9]+)_altseq[0-9]*_([A-Za-z]+)_to_([A-Za-z]+)$",
      "\\1:\\2:\\3:\\4"
    )
  )

maptlabeled <- maptlabeled %>% select(pos, base_ref, base_alt, adj.P.Val, logFC)
#maptlabeled <- maptlabeled %>% rename("BCALM_adjPVal" = "adj.P.Val")
#maptlabeled <- maptlabeled %>% rename("BCALM_logFC" = "logFC")

promai <- promai %>% select(pos, ref, alt, score)
promai <- promai %>% dplyr::rename("base_ref" = "ref")
promai <- promai %>% rename("base_alt" = "alt")
promai <- promai %>% rename("PromAIScore" = "score")

alphagenomeprom <- alphagenomeprom %>% select(pos, base_ref, base_alt, raw_score)
alphagenomeprom <- alphagenomeprom %>% rename("AlphaGenome_rawScore" = "raw_score")

merged_up <- overlapup %>%
  inner_join(maptlabeled, by = c("pos", "base_alt", "base_ref")) %>%
  inner_join(promai, by = c("pos", "base_alt", "base_ref")) %>%
  inner_join(alphagenomeprom, by = c("pos", "base_alt", "base_ref"))
merged_down <- overlapdown %>% 
  inner_join(maptlabeled, by = c("pos", "base_alt", "base_ref")) %>%
  inner_join(promai, by = c("pos", "base_alt", "base_ref")) %>%
  inner_join(alphagenomeprom, by = c("pos", "base_alt", "base_ref"))
write_tsv(merged_up, "/cluster/home/rhauser/files/rmhex03/AIfromHenry/merged_allup.tsv")

write_tsv(merged_down, "/cluster/home/rhauser/files/rmhex03/AIfromHenry/merged_alldown.tsv")

merged_all <- rbind(merged_up, merged_down)

library(htmlwidgets)
library(plotly)
library(ggrepel)
merged_all <- merged_all %>%
  mutate(snp = paste("chr17", pos, base_ref, base_alt, sep = ":"))
merged_all <- merged_all %>% arrange(by=abs(BCALM_logFC))


compare <- ggplot(merged_all, aes(x=PromAIScore, y=AlphaGenome_rawScore, color = BCALM_logFC, text=snp)) +
  geom_point() +
  theme_classic() +
  scale_color_gradient2(low = "dodgerblue", mid = "white", high = "tomato", midpoint = 0, limits = c(-0.65, 0.65))+
  coord_cartesian(xlim = c(-0.9,0.9), ylim=c(-0.17,0.17)) 

ggsave(compare, file="alphagenome_promai_bcalm_directionagree.pdf", height=4, width=5)
    

compare_p <- ggplotly(compare, tooltip="text")

htmlwidgets::saveWidget(compare_p, file = "alphagenome_promai_bcalm_directionagree.html", selfcontained = TRUE)


merged_all <- maptlabeled %>% inner_join(alphagenomeprom, by = c("pos", "base_ref", "base_alt"))
merged_all <- merged_all %>% inner_join(promai, by = c("pos", "base_ref", "base_alt"))

merged_all <- merged_all %>%
  mutate(
    group = case_when(
      # --- PromAIScore == 0 cases ---------------------------------------------
      PromAIScore == 0 & (
        (logFC > 0  & AlphaGenome_rawScore > 0) |
        (logFC < 0  & AlphaGenome_rawScore < 0)
      ) ~ "one_agree",

      PromAIScore == 0 & (
        (logFC > 0 & AlphaGenome_rawScore < 0) |
        (logFC < 0 & AlphaGenome_rawScore > 0)
      ) ~ "no_agree",

      # --- Regular cases (nonzero PromAIScore) --------------------------------
      (logFC > 0 & AlphaGenome_rawScore < 0 & PromAIScore < 0) |
      (logFC < 0 & AlphaGenome_rawScore > 0 & PromAIScore > 0) ~ "no_agree",

      (logFC > 0 & AlphaGenome_rawScore > 0 & PromAIScore > 0) |
      (logFC < 0 & AlphaGenome_rawScore < 0 & PromAIScore < 0) ~ "all_agree",

      (logFC > 0 & ((AlphaGenome_rawScore > 0 & PromAIScore < 0) |
                    (AlphaGenome_rawScore < 0 & PromAIScore > 0))) |
      (logFC < 0 & ((AlphaGenome_rawScore < 0 & PromAIScore > 0) |
                    (AlphaGenome_rawScore > 0 & PromAIScore < 0))) ~ "one_agree",

      TRUE ~ "missing"
    )
  )
merged_all <- merged_all %>% mutate(
    sig = case_when(
      adj.P.Val <= 0.05 ~ "sig",
      TRUE ~ "ns"))

merged_all  <- merged_all  %>%
  mutate(
    snp = paste0("chr17:", pos, ":", base_ref, ":", base_alt)
  )
merged_all <- merged_all %>% filter(!is.na(logFC))
merged_all <- merged_all %>%
  mutate(group = factor(group, levels = c("all_agree", "one_agree", "no_agree")))
merged_all <- merged_all %>%
    mutate(
    region = case_when(
      pos %in% c(45894271, 45894272, 45894273, 45894275, 45894277, 45894278,
                 45894279, 45894288, 45894289, 45894290, 45894291, 45894293,
                 45894294, 45894295, 45894296, 45894297) ~ "down",
      pos %in% c(45894685, 45894686, 45894687, 45894688, 45894689,
                 45894690, 45894691) ~ "up",
      TRUE ~ "other"
    )
  )
merged_all <- merged_all %>%
  mutate(region = factor(region, levels = c("down", "up", "other")))

compare <- ggplot(merged_all, aes(x = PromAIScore,
                                  y = AlphaGenome_rawScore,
                                  text = snp)) +
  # Base layer: non-significant points (gray)
  geom_point(
    data = subset(merged_all, sig == "ns"),
    color = "gray",
    size = 2) +
  # Significant points: logFC fill (plasma), outline by region
  geom_point(
    data = subset(merged_all, sig == "sig"),
    aes(color = logFC),
    size = 2)+
  # Fill gradient by logFC
  scale_color_viridis_c(option = "plasma") +
  theme_classic() +
  facet_wrap(~group)

ggsave(compare, file="alphagenome_promai_bcalm_all.pdf", height=4, width=11)
library(plotly)
compare_p <- ggplotly(compare, tooltip="text")
htmlwidgets::saveWidget(compare_p, file = "alphagenome_promai_bcalm_all.html", selfcontained = TRUE)


compare <- ggplot(merged_all, aes(x = PromAIScore,
                                  y = AlphaGenome_rawScore,
                                  text = snp)) +
  # Base layer: non-significant points (gray)
  geom_point(
    data = subset(merged_all, sig == "ns"),
    color = "gray",
    size = 2) +
  # Significant points: logFC fill (plasma), outline by region
  geom_point(
    data = subset(merged_all, sig == "sig"),
    aes(color = region),
    size = 2)+
  # Fill gradient by logFC
  scale_color_manual(values=c(down = "dodgerblue", up="tomato", other="#261823")) +
  theme_classic() +
  facet_wrap(~group)

#ggsave(compare, file="alphagenome_promai_bcalm_all_region.pdf", height=4, width=11)

install.packages("ggpubr")
library(ggpubr)

compare <- ggplot(merged_all, aes(x=PromAIScore, y=logFC)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label")
ggsave(compare, file="promaiVsMPRA.pdf", height=4, width=4)

res <- cor.test(merged_all$PromAIScore, merged_all$logFC, method = "spearman", exact =FALSE)

res$estimate  # Spearmans's rho = 0.04819852
res$p.value   # P-value 0.0001892385

compare <- ggplot(merged_all, aes(x=AlphaGenome_rawScore, y=logFC)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label")
ggsave(compare, file="alphaGenomeVsMPRA.pdf", height=4, width=4)

res <- cor.test(merged_all$AlphaGenome_rawScore, merged_all$logFC, method = "spearman", exact = FALSE)

res$estimate  # Spearmans's rho 0.02767953
res$p.value   # P-value 0.03211802

compare <- ggplot(merged_all%>%filter(adj.P.Val<=0.05), aes(x=PromAIScore, y=logFC)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label")
ggsave(compare, file="promaiVsMPRAonlysig.pdf", height=4, width=4)

sig_data <- merged_all %>% dplyr::filter(adj.P.Val <= 0.05)

res <- cor.test(sig_data$PromAIScore, sig_data$logFC, method = "spearman", exact = FALSE)

r<-res$estimate  # Spearmans's rho r 0.2923967 
res$p.value   # p-value 1.102299e-09

compare <- ggplot(merged_all%>%filter(adj.P.Val<=0.05), aes(x=AlphaGenome_rawScore, y=logFC)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label")
ggsave(compare, file="AlphaGenomeVsMPRAonlysig.pdf", height=4, width=4)

sig_data <- merged_all %>% dplyr::filter(adj.P.Val <= 0.05)

res <- cor.test(sig_data$AlphaGenome_rawScore, sig_data$logFC, method = "spearman", exact = FALSE)

res$estimate  # Spearmans's rho 0.1622406
res$p.value   # p-value 0.0008710601

compare <- ggplot(merged_all, aes(x=AlphaGenome_rawScore, y=PromAIScore)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label")
ggsave(compare, file="AlphaGenomeVsPromAI.pdf", height=4, width=4)

res <- cor.test(merged_all$AlphaGenome_rawScore, merged_all$PromAIScore, method = "spearman", exact = FALSE)

res$estimate  # Spearmans's rho 0.153362  
#rsquare <- r^2 #0.1288706 
res$p.value   # p-value 7.177675e-33
```

#redo of motifbreakr graphs with new seq
```{r}

mymotifs <- readRDS("/cluster/home/rhauser/files/rmhex03/motifbreakr/motifbreakr_SNVs_results.rds")
library(motifbreakR)
library(BSgenome.Hsapiens.UCSC.hg38)

mymotifs_tib <- as_tibble(mymotifs)
write_tsv(mymotifs_tib, "/cluster/home/rhauser/files/rmhex03/motifbreakr/motifbreakr_SNVs_results.tsv")

strongmotifs <- mymotifs[mymotifs$effect == "strong"]
mypromoter <- GRanges(seqnames="chr17", ranges=IRanges(start=c(45894158), end=c(45894411)))

strongprommotifs <- strongmotifs[strongmotifs %over% mypromoter]
strongprommotifsneg <- strongprommotifs[strongprommotifs$alleleEffectSize < 0]

pdf(file="chr17_45894294_G_A.pdf", width = 4.44, height = 10, onefile=FALSE)
maptmotif1 <- plotMB(strongprommotifsneg, rsid = "chr17:45894294:G:A", effect="strong")
dev.off()


strongprom_genes_counts <- table(mcols(strongprommotifs)$geneSymbol) %>% sort()

strongprommotifs_tib <- as_tibble(strongprommotifs)
mymotifs_tib <- as_tibble(mymotifs)

maptlabeled$chr <- "chr17"
maptlabeled_snvid <- maptlabeled %>%
  mutate(snvid = paste0(chr, ":", pos, ":", base_ref, ":", base_alt))

#filter for only a negative motif breaker effect
strongprommotifs_tibdown <- strongprommotifs_tib %>% filter(alleleEffectSize < 0)
mymotifs_tibdown <- mymotifs_tib %>% filter(alleleEffectSize < 0)

#filter for only in loss of activity region
maptlabeled_snvidfilt <- maptlabeled_snvid %>% filter(between(pos, 45894271, 45894297)) %>% arrange(by=pos)
maptlabeled_snvidfilt <- maptlabeled_snvidfilt %>% filter(adj.P.Val <= 0.05)
maptlabeled_snvidfilt <- maptlabeled_snvidfilt %>% filter(logFC <0)
strongprommotifs_tibdown_filtered <- strongprommotifs_tibdown %>%
  filter(SNP_id %in% maptlabeled_snvidfilt$snvid)
strongprommotifs_tibdown_filtered_count <- strongprommotifs_tibdown_filtered %>%
  count(geneSymbol, name = "count")
mymotifs_filtered <- mymotifs_tibdown %>%
  filter(SNP_id %in% maptlabeled_snvidfilt$snvid)


mymotifs_filtered_count <- mymotifs_filtered  %>%
  count(geneSymbol, name = "count")


#get motif group names
motifgroups <- read_tsv("/cluster/home/aanderson/myers/scAnalysis/JASPAR/clusters_motif_names.tab", col_names=c("group", "TFlist"))
motifgroups <- motifgroups %>%
  mutate(TFlist = as.character(TFlist))
motifgroups2 <-motifgroups %>%  separate_rows(TFlist, sep = ",") %>%
  mutate(TFlist = gsub("::", "_", TFlist))
motifgroups2 <- motifgroups2 %>% dplyr::rename(TFgroup = group)
motifgroups2 <- motifgroups2 %>% dplyr::rename(geneSymbol = TFlist)


#add motifgroups
motifgroups2_aggregated <- motifgroups2 %>%
  group_by(geneSymbol) %>%
  summarise(TFgroup = paste(unique(TFgroup), collapse = ", "), .groups = "drop")
prommotifsjoin <- left_join(strongprommotifs_tibdown_filtered_count, motifgroups2_aggregated, by = "geneSymbol")

strongprommotifs_tibdown_filtered2 <- left_join(strongprommotifs_tibdown_filtered, motifgroups2_aggregated, by="geneSymbol")
write_tsv(strongprommotifs_tibdown_filtered2, "/cluster/home/rhauser/MotifbreakerResults.tsv")

tfbar <- ggplot(prommotifsjoin, aes(x=count, y=reorder(geneSymbol, count), fill=TFgroup)) +
  geom_col() +
  theme_classic() +
  labs(y="Transcription Factor Binding Sites")
ggsave(filename="/cluster/home/rhauser/files/graphs/paperfigures/newvariantplots/TFs_bcalm_down_strongnegeffect_onlydownregion_newseq.pdf", plot=tfbar, width=5.44, height=5.44, units="in")

prommotifsjoin_alldown <- left_join(mymotifs_filtered_count, motifgroups2_aggregated, by = "geneSymbol")
tfbar <- ggplot(prommotifsjoin_alldown, aes(x=count, y=reorder(geneSymbol, count), fill=TFgroup)) +
  geom_col() +
  theme_classic() +
  labs(y="Transcription Factor Binding Sites")
ggsave(filename="/cluster/home/rhauser/files/graphs/paperfigures/newvariantplots/TFs_bcalm_down_strongandweaknegeffect_onlydownregion_newseq.pdf", plot=tfbar, width=5.44, height=5.44, units="in")


tfbar <- ggplot(prommotifsjoin, aes(x=count, y=reorder(geneSymbol, count), fill=TFgroup)) +
  geom_col() +
  theme_classic() +
  scale_fill_manual(values=c("#6acfdc", "#ec5bb4")) +
  labs(y="Transcription Factor Binding Sites")
ggsave(filename="/cluster/home/rhauser/files/graphs/paperfigures/newvariantplots/TFs_bcalm_down_strongnegeffect_onlydownregion_newseq_newcolors.pdf", plot=tfbar, width=5, height=5.44, units="in")


```



#spliceAI scores
```{r}
spliceai <- read_tsv("/cluster/home/rhauser/files/rmhex03/spliceAI/spliceAIscores_promupregion.tsv")
spliceai_donor <- spliceai %>% select(POS, REF, ALT, DonorLoss, DonorGain)
spliceai_donor_long <- spliceai_donor %>% pivot_longer(cols = c("DonorLoss", "DonorGain"), names_to="Type", values_to = "Score")


spliceaifig <- ggplot(spliceai_donor_long , aes(fill=ALT, y=Score, x=POS)) +
  geom_bar(position="dodge", stat="identity", width=0.7) +
  theme_minimal() +
  scale_fill_manual(
    values = c(
      "A" = "#1b998b", 
      "T" = "#ff6347",  
      "C" = "#5cadff", 
      "G" = "#fecb34"
    )) +
  geom_hline(yintercept=0) +
  facet_wrap(~Type)
ggsave(filename="spliceAIfig.pdf", plot=spliceaifig, width=15, height=4.44, units="in")
```

#rs11575896
```{r}
ancestry <- read_tsv("/cluster/home/rhauser/files/rmhex03/variantinfo/frq_outputs/summary.txt")
ancestry_long <- ancestry %>% pivot_longer(cols = c("AlleleFreq_AD", "AlleleFreq_nonAD"), names_to="Group", values_to = "AlleleFreq")

#order factors
ancestry_long <- ancestry_long %>%
  mutate(
    Ancestry = factor(Ancestry, levels = c("Total", setdiff(unique(Ancestry), "Total"))),
    Group = factor(Group, levels = c("AlleleFreq_nonAD", "AlleleFreq_AD"))
  )


ancestryp <- ggplot(ancestry_long, aes(fill=Group, x=Ancestry, y=AlleleFreq)) +
  geom_bar(position="dodge", stat="identity", width=0.7) +
  theme_minimal() +
  scale_fill_manual(
    values = c(
      "AlleleFreq_AD" = "#98e6ec",
      "AlleleFreq_nonAD" = "#e0e0e0"
    )
  )
ggsave(filename="rs11575896ancestry.pdf", plot=ancestryp, width=15, height=4.44, units = "in")

ancestry_long_nototal <- ancestry_long %>% filter(Ancestry != "Total")
ancestry_long_nototal <- ancestry_long_nototal %>% pivot_longer(cols = c("n_AD", "n_nonAD"), names_to="Group2", values_to = "Count")

ancestryp2 <- ggplot(ancestry_long_nototal, aes(x=Group2, y=Count, fill = Ancestry)) + 
  geom_bar(position="stack", stat="identity", show.legend = TRUE) +
  theme_minimal() + 
  ylab("Count") + 
  ggtitle("ADSP AD Counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(filename="ADSPancestry.pdf", plot=ancestryp2, width=4, height=4.44, units="in")


```

#acetylation and conservation
```{r}

library(tidyverse)
library(ggplot2)
library(plotgardener)
library(GenomicRanges)
library(rtracklayer)

histoneAc <- import.bw("/cluster/home/jloupe/Outputs_sorted_cromwell/Sep2021_ChIPseq/H3K27ac_DLPFC-NEUN_Output/signal/pooled-rep/basename_prefix.pooled_x_basename_prefix.pooled.pval.signal.bigwig")
histoneAc17 <- histoneAc[seqnames(histoneAc) == "chr17"]

prom <- GRanges("chr17:45893336-45895735")
histoneAcProm <- subsetByOverlaps(histoneAc, prom)
histoneAcPromTib <- as_tibble(histoneAcProm)

cons <- import.bw("/cluster/home/aanderson/cross_species/external_data/cactus241way.phyloP.bw", which = GRanges("chr17", IRanges(1, seqlengths(BigWigFile("/cluster/home/aanderson/cross_species/external_data/cactus241way.phyloP.bw"))["chr17"]))
)

#cap signal (from paperfigers.Rmd)
avg_neurons <- read_tsv("/cluster/home/rhauser/files/rmhex02/onesidedneurons3repswithcontrolsneglog10pval.sorted.mean.bed", col_names=c("chr", "start", "end", "neglog10pval"))
avg_neurons$start <- avg_neurons$start + 1
avg_neurons$score <- avg_neurons$neglog10pval

avg_neurons <- avg_neurons %>% arrange(by=start)
cappedsignal <- avg_neurons %>% mutate(score=pmin(score, 15))
```


####mapt prom region plotgardener
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="maptprom_onetailedp05_sigline_acandcons.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45893536, chromend = 45895535, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")

 
#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#plot conservation
consplot <- plotSignal(data = cons, params = params_a, y = 0.2, height = 0.5, fill = "#d88c9a", linecolor="#d88c9a")
annoYaxis(plot = consplot, axisLine = TRUE, fontsize = 8, at=c(-10,9))

#plot acetylation
acplot <- plotSignal(data =histoneAcProm, params = params_a, y = 0.9, height = .5, fill = "#8E7DBE", linecolor="#8E7DBE")
annoYaxis(plot = acplot, axisLine = TRUE, fontsize = 8, at=c(0,91))


threshold = -log10(0.05)



#plot ompra
oligologplot <- plotSignal(data = cappedsignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, at=c(0,15), fontsize = 8, digits=3)

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="region5_onetailedp05_sigline_acandcons.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45821000, chromend = 45823000, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")

 
#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#plot conservation
consplot <- plotSignal(data = cons, params = params_a, y = 0.2, height = 0.5, fill = "#d88c9a", linecolor="#d88c9a")
annoYaxis(plot = consplot, axisLine = TRUE, fontsize = 8, at=c(-12,9))

#plot acetylation
acplot <- plotSignal(data =histoneAc17, params = params_a, y = 0.9, height = .5, fill = "#8E7DBE", linecolor="#8E7DBE")
annoYaxis(plot = acplot, axisLine = TRUE, fontsize = 8,at=c(0, 5.5))


threshold = -log10(0.05)



#plot ompra
oligologplot <- plotSignal(data = cappedsignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, fontsize = 8, digits=3, at=c(0,4.5))

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####bri region 45971424_45972091
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="plotgardenerbri45971424_45972091_acandcons.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45971424, chromend = 45972091, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")

 
#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#plot conservation
consplot <- plotSignal(data = cons, params = params_a, y = 0.2, height = 0.5, fill = "#d88c9a", linecolor="#d88c9a")
annoYaxis(plot = consplot, axisLine = TRUE, fontsize = 8, at=c(-8,9))

#plot acetylation
acplot <- plotSignal(data =histoneAc17, params = params_a, y = 0.9, height = .5, fill = "#8E7DBE", linecolor="#8E7DBE")
annoYaxis(plot = acplot, axisLine = TRUE, fontsize = 8, at=c(0,8.5))


threshold = -log10(0.05)



#plot ompra
oligologplot <- plotSignal(data = cappedsignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, fontsize = 8, digits=3, at=c(0,3))

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```


####region 18
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="region18_acandcons.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 45779500, chromend = 45782500, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")

 
#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#plot conservation
consplot <- plotSignal(data = cons, params = params_a, y = 0.2, height = 0.5, fill = "#d88c9a", linecolor="#d88c9a")
annoYaxis(plot = consplot, axisLine = TRUE, fontsize = 8, at=c(-7,7.5))

#plot acetylation
acplot <- plotSignal(data =histoneAc17, params = params_a, y = 0.9, height = .5, fill = "#8E7DBE", linecolor="#8E7DBE")
annoYaxis(plot = acplot, axisLine = TRUE, fontsize = 8,at=c(0, 5))


threshold = -log10(0.05)



#plot ompra
oligologplot <- plotSignal(data = cappedsignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, fontsize = 8, digits=3,at=c(0,3))

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####region 8
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="region8_acandcons.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 46029900, chromend = 46031100, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")

 
#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#plot conservation
consplot <- plotSignal(data = cons, params = params_a, y = 0.2, height = 0.5, fill = "#d88c9a", linecolor="#d88c9a")
annoYaxis(plot = consplot, axisLine = TRUE, fontsize = 8, at=c(-8.5,9))

#plot acetylation
acplot <- plotSignal(data =histoneAc17, params = params_a, y = 0.9, height = .5, fill = "#8E7DBE", linecolor="#8E7DBE")
annoYaxis(plot = acplot, axisLine = TRUE, fontsize = 8, at=c(0, 12))


threshold = -log10(0.05)



#plot ompra
oligologplot <- plotSignal(data = cappedsignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, fontsize = 8, digits=3, at=c(0,9.5))

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```

####region 9
```{r}
#un-comment to save as a pdf in working directory labeled Rplots.pdf
pdf(file="region9_acandcons.pdf", width = 8, height = 4.25, onefile=FALSE)

## Create a plotgardener page
pageCreate(width = 8, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

pageGuideHide()

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 46036535, chromend = 46039410, assembly = "hg38", x = 1.5, width =5, just = c("left", "top"), default.units = "inches")

 
#plot variation
#plotRanges(data = variants1bed, params = params_a, order. ="random", y = .5, height = .3, collapse = TRUE, fill = "#F433AB")

#plot conservation
consplot <- plotSignal(data = cons, params = params_a, y = 0.2, height = 0.5, fill = "#d88c9a", linecolor="#d88c9a")
annoYaxis(plot = consplot, axisLine = TRUE, fontsize = 8, at=c(-12.5, 9))

#plot acetylation
acplot <- plotSignal(data =histoneAc17, params = params_a, y = 0.9, height = .5, fill = "#8E7DBE", linecolor="#8E7DBE")
annoYaxis(plot = acplot, axisLine = TRUE, fontsize = 8, at=c(0,25.5))


threshold = -log10(0.05)



#plot ompra
oligologplot <- plotSignal(data = cappedsignal, params = params_a, negData = TRUE, y = 1.75, height = 0.75, fill = "#86EAE6", line = "#86EAE6")
annoSegments(
  x0=unit(0, "npc"), y0=threshold,
  x1=unit(1, "npc"), y1=threshold,
  plot=oligologplot, default.units="native",
  linecolor = "grey40", lty=2
)
annoYaxis(plot = oligologplot,axisLine = TRUE, fontsize = 8, digits=3, at=c(0,6))

#plot cCres
#plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = "#F2D0A9")
plotRanges(data = encodecres, params = params_a, order. ="width", collapse=TRUE, y = 2.75, height = .075, fill = encodecres$hexcolor)


geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("black"))

#making gene plot
plotGenes(params = params_a, stroke =0.5, fontsize = 8, strandLabels = TRUE, y = 3, height =1, geneHighlights = geneHighlights, geneBackground = "black")

plotGenomeLabel(params = params_a, y =4, length = 5, scale = "bp", commas=TRUE)

dev.off()


```



#conservation graphing zoomins
```{r}
cons_tib <- as_tibble(cons)
cons_tib_loss <- cons_tib %>% filter(between(start, 45894270, 45894299))

maptactive <- ggplot(cons_tib_loss, aes(y=score, x=start)) +
  geom_bar(stat="identity", fill="orange", color = "orange") +
  theme_minimal() +
  geom_hline(yintercept=0) +
  scale_y_continuous(
    limits = c(-5, 10),
    breaks = seq(-5, 10),
    expand = c(0, 0)
  )
ggsave(filename="maptbcalm_cons_30bases_losactivity.pdf", plot=maptactive, width=15, height=4.44, units="in")


cons_tib_active <- cons_tib %>% filter(between(start, 45894673, 45894702))

maptactive <- ggplot(cons_tib_active, aes(y=score, x=start)) +
  geom_bar(stat="identity", fill="orange", color = "orange") +
  theme_minimal() +
  geom_hline(yintercept=0) +
  scale_y_continuous(
    limits = c(-1.5, 9),
    breaks = seq(-1.5, 9),
    expand = c(0, 0)
  )
ggsave(filename="maptbcalm_cons_30bases_gainactivity.pdf", plot=maptactive, width=15, height=4.44, units="in")

```


#plot overview figure
```{r}
library(tidyverse)
library(ggplot2)
library(GenomicRanges)
library(plotgardener)

#all bacs
allbacs <- read_tsv("/cluster/home/rhauser/files/rmhex01/bacslabel_original.bed", col_names=c("chr", "start", "end", "name"))

#baccoverage
cov <- read_tsv(file="/cluster/home/rhauser/files/from_nick/OutFiles.chr17.depth.bedGraph", col_names=c("chr", "start", "end", "score"))

cov$end <- cov$end - 1
cov$width <- cov$end - cov$start
cov$strand <- '*'
cov <- cov %>% dplyr::select(chr, start, end, width, strand, score)

rnaseqregion <- GRanges(
  seqnames = "chr17",
  ranges = IRanges(44705302, 46838000)
)

inversionregion <- GRanges(
  seqnames = "chr17",
  ranges = IRanges(45307498,46836265)
)

#create ideogramtrack
itrack <- IdeogramTrack(genome = "hg38", chromosome = "chr17")
highlightTrack <- AnnotationTrack(
  rnaseqregion,
  genome = "hg38",
  name = "BAC coverage",
  fill = "red",
  col = "red"
)

ht <- HighlightTrack(
  trackList = list(highlightTrack),
  start = start(rnaseqregion),
  end = end(rnaseqregion)
)

pdf(file="chr17ideogram.pdf", width = 8, height = 4.25, onefile=FALSE)
plotTracks(
  list(itrack, ht),
  chromosome = "chr17",
  main = "Chromosome 17 (hg38) — highlighted region"
)
dev.off()


mycrisprtargets <- GRanges(
  seqnames = "chr17",
  ranges= IRanges(
    start = c(44905302, 45371797, 45803209, 45821846, 45837167, 45855626, 46030475, 46037972, 46072943, 45650602, 45828414, 45655341, 45819754, 45781206, 45893945),
    end = c(44905302, 45371797, 45803209, 45821846, 45837167, 45855626, 46030475, 46037972, 46072943, 45650602, 45828414, 45655341, 45819754, 45781206, 45893945)
))
mycrisprtargets_tib <- as_tibble(mycrisprtargets)



pdf(file="regionoverview_new.pdf", width = 15, height = 4.25, onefile=FALSE)
## Create a plotgardener page
pageCreate(width = 15, height = 4.25, default.units = "inches", showGuides=FALSE, bg=NA)

#set general parameters
params_a <- pgParams(chrom = "chr17", chromstart = 44705300, chromend = 46172143, assembly = "hg38", x = 1.5, width =12, just = c("left", "top"), default.units = "inches")

#plot inversion
#plotRanges(
#  data = inversionregion, params = params_a, order. ="random", y = 0.7, height = .75, boxHeight = 0.15, fill = "black" 
#)

#making gene plot, highlighting MAPT, and label
geneHighlights <- data.frame("geneName" = c("MAPT"), "color" = c("#f56d38"))

plotGenes(params = params_a, stroke =0.75, fontsize =10, strandLabels = TRUE, y = 2, height =1, geneHighlights = geneHighlights, geneBackground ="#cbcbcb")

plotGenomeLabel(params = params_a, y =3.5, length = 12, scale = "Mb")

dev.off()




targets <- c(44905302, 45371797, 45803209, 45821846, 45837167, 45855626, 46030475, 46037972, 46072943, 45650602, 45828414, 45655341, 45819754, 45781206, 45893945)

mycrisprtargets_tib <- mycrisprtargets_tib %>% arrange(by=start)
mycrisprtargets_tib$plotheight <- c(0, 0, 0, 0.1, 0, 0, 0.1, 0.2, 0.3, 0, 0, 0, 0, 0.1, 0)
mycrisprtargets_tib$plotheight2 <- 0


crisptar <- ggplot(mycrisprtargets_tib, aes(x=start, y=plotheight)) +
  geom_point(color="#86EAE6", size=3, shape="|") +
  theme_minimal() +
  coord_cartesian(xlim = c(44705300, 46172143), ylim=c(-0.5,0.5)) +
  scale_x_continuous(breaks=c(44705300, 46172143), labels=c("44902000", "46172143"))
ggsave(filename="crisprguidelocs_new.pdf", plot=crisptar, width=10, height=2.22, units="in")


```
#corr mapt promoter cell type
```{r}
mapt <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/neurons/bcalm_mapt_neurons_10bcthres.tsv")

mapt <- mapt %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < 0 ~ "sig_down",
      adj.P.Val > 0.05 ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

mapt <- mapt %>% mutate(position = as.numeric(str_extract(element, "(?<=maptprom_chr17_)[0-9]+")))

#implementing a logFC thres of 0.1
mapt <- mapt %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0.1 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < -0.1 ~ "sig_down",
      adj.P.Val > 0.05 ~ "ns",
      adj.P.Val <= 0.05 & logFC >= -0.1 & logFC <= 0.1 ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))


# Make sure log2FoldChange is numeric
mapt$logFC <- as.numeric(mapt$logFC)

mapt$sig2 <- as.character(mapt$sig2)

maptheks <- read_tsv("/cluster/home/rhauser/files/rmhex03/MPRAsnakeflowcount/bcalm/heks/bcalm_mapt_heks_10bcsthres.tsv")
maptheks <- maptheks %>% mutate(
    sig2 = case_when(
      adj.P.Val <= 0.05 & logFC > 0.1 ~ "sig_up",
      adj.P.Val <= 0.05 & logFC < -0.1 ~ "sig_down",
      adj.P.Val > 0.05 ~ "ns",
      adj.P.Val <= 0.05 & logFC >= -0.1 & logFC <= 0.1 ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

maptheks <- maptheks %>% mutate(position = as.numeric(str_extract(element, "(?<=maptprom_chr17_)[0-9]+")))
maptheks$sig2 <- as.character(maptheks$sig2)

maptheks_renamed <- maptheks %>%
  rename_with(~ paste0(.x, ".heks"), .cols = -c(position, allele))

maptcombined <- mapt %>%
  inner_join(maptheks_renamed, by = c("position", "allele"))

maptcombined <- maptcombined %>% mutate(
    shared = case_when(
      sig2 == "sig_up" & sig2.heks == "sig_up" ~ "both",
      sig2 == "sig_down" & sig2.heks == "sig_down" ~ "both",
      sig2 == "sig_down" & sig2.heks == "sig_up" ~ "both",
      sig2 == "sig_up" & sig2.heks == "sig_down" ~ "both",
      sig2 == "sig_up" & sig2.heks == "ns" ~ "signeurons",
      sig2 == "sig_down" & sig2.heks == "ns" ~ "signeurons",
      sig2 == "ns" & sig2.heks == "sig_up" ~ "sigheks",
      sig2 == "ns" & sig2.heks == "sig_down" ~ "sigheks",
      sig2 == "ns" & sig2.heks == "ns" ~ "ns",
      TRUE    ~ NA_character_  # Fallback if none are TRUE
    ))

maptcombined$shared <- as.character(maptcombined$shared)
maptcombined <- maptcombined %>% filter(!is.na(shared))


maptcombined$shared <- factor(
  maptcombined$shared,
  levels = c("ns", "sigheks", "signeurons", "both")
)

maptcombined <- maptcombined %>% arrange(shared)
write_tsv(maptcombined, "maptcombined.tsv")

maptcellcorr <- ggplot(maptcombined, aes(x = logFC.heks, y = logFC, color = shared)) +
  geom_point(size = 1.5) +
  scale_color_manual(name="Adj.P.Val & logFC cutoff",
    values = c(
      "both"   = "#8332b5",
      "sigheks"    = "#6acfdc",
      "signeurons" = "#ec5bb4",
      "ns"        = "#e0e0e0"
    ),
    labels = c(
    "ns"        = "None",
    "sigheks"    = "HEKs",
    "signeurons" = "Neurons",
    "both"   = "HEKs & Neurons"
  )) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") +  
  geom_hline(yintercept = 0, linetype = "dotted", color = "black") +
  labs(
    x = "HEK293FT logFC",
    y = "Neuron logFC",
    title = "MAPT Promoter Mutagenesis \nNeurons vs. HEK293FT"
  ) +
  theme_classic() +
  theme(legend.position = "right") # +
#  scale_y_continuous(limits=c(-1.2,0.8), breaks = seq(-1.2,0.8, by =0.2), labels=scales::label_number(accuracy=0.1)) +
#  scale_x_continuous(limits=c(-0.8,0.8), breaks=seq(-0.8,0.8, by=0.2), labels=scales::label_number(accuracy=0.1))
ggsave(filename ="maptpromcorr_logFCcutoff_10bcthres.pdf", plot=maptcellcorr, width=5.44, height=4.44, units="in")
res <- cor.test(maptcombined$logFC, maptcombined$logFC.heks, method = "spearman", exact =FALSE)

res$estimate  # Spearmans's rho = 0.336217
res$p.value   # P-value 1.556827e-210

maptcombined_nodel <- maptcombined %>% filter(allele != "E")

```
